---
title: "BTN Cocktail Inference Pipeline"

author: "This RMarkdown includes all analyses resulting from the computational methods developed in <insert name of Paper and authors> to infer the BTN Cocktail.  Written by Prakriti Paul Chacha"
  
output:
  github_document:
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Import Relevant Libraries
> Please install any missing packages.

```{r eval=FALSE, error=FALSE, warning=FALSE}
library(URD)
library(Seurat)
library(ggplot2)
library(tidyverse)
library(stats)
library(Matrix)
library(stringr)
library(edgeR)
library(rlang)
library(grpreg)
library(glmnet)
library(dplyr)
library(purrr)
library(stringr)
library(reticulate)
source("./helper_files/final_main_pipeline_start_script.R")
source("./helper_files/final_helper_functions.R")
```
# (1) Identify Precursor, Mother, and their Sibling(s) Segment Cell Populations on Virtual Neural Lineage Tree
Segment cells refers to the cells along the tree segment/branch that will contain
Precursor, Precursor Sibling, and Mother pseudostates.

Note: "CNS.k300s6w4.tree.built.name" is the URD object generated from time-course
      single-cell RNA-seq data in Chen et. al* for the Central Nervous System (CNS)
      of Ciona intestinalis.
      It contains cell names, their pseudotimes, the Virtual Lineage Tree, and all
      URD object attributes.

*Cao, Chen, et al. "Comprehensive single-cell transcriptome lineages of a proto-vertebrate." Nature 571.7765 (2019): 349-354.

## Precursor Segment 

### Get IDs of BTN Precursor Segment and its Sibling Segment
```{r eval=FALSE}
# Get ID from URD Object.
BTN_ID <- "36"
BTN_sibling_ID <- segSiblings(CNS.k300s6w4.tree.built.name, BTN_ID, include.self = F)
```

### Get BTN Precursor and Sibling Segment Cells
```{r eval=FALSE}
# 463
BTN_precursor_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", BTN_ID)
# 575
BTN_sibling_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", BTN_sibling_ID)
```

### Order Cells by Pseudotime
```{r eval=FALSE}
pseudotimed_BTN_precursor_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, BTN_ID, length(BTN_precursor_segment_cells), ascending=TRUE)

pseudotimed_BTN_precursor_sibling_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, BTN_sibling_ID, length(BTN_sibling_segment_cells), ascending=TRUE)
```

## Mother Segment

### Get IDs of BTN Mother Segment and its Siblings
```{r eval=FALSE}
BTN_mother_ID <- "70"
# 63, 73
BTN_mother_sibling_IDs <- segSiblings(CNS.k300s6w4.tree.built.name, BTN_mother_ID, include.self = F)

BTN_mother_sibling_63_ID <- BTN_mother_sibling_IDs[1]
BTN_mother_sibling_73_ID <- BTN_mother_sibling_IDs[2]
```

### Get BTN Mother and Sibling Segment Cells
```{r eval=FALSE}
# 463
BTN_mother_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", BTN_mother_ID)
# 853
BTN_mother_sibling_63_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", BTN_mother_sibling_63_ID)
# 25
BTN_mother_sibling_73_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", BTN_mother_sibling_73_ID)
```

### Order Cells by Pseudotime
```{r eval=FALSE}
pseudotimed_BTN_mother_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, BTN_mother_ID, length(BTN_mother_segment_cells), ascending=FALSE)

pseudotimed_BTN_mother_sibling_63_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, BTN_mother_sibling_63_ID, length(BTN_mother_sibling_63_segment_cells), ascending=FALSE)

pseudotimed_BTN_mother_sibling_73_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, BTN_mother_sibling_73_ID, length(BTN_mother_sibling_73_segment_cells), ascending=FALSE)
```

## Visualize Cells on Virtual Neural Lineage Tree
```{r eval=FALSE}
BTN_segment_cells_list <- list("pseudotimed_BTN_precursor_segment_cells" = pseudotimed_BTN_precursor_segment_cells,
                               "pseudotimed_BTN_precursor_segment_cells" = pseudotimed_BTN_precursor_segment_cells,
                               "pseudotimed_BTN_mother_segment_cells" = pseudotimed_BTN_mother_segment_cells,
                               "pseudotimed_BTN_mother_sibling_63_segment_cells" = pseudotimed_BTN_mother_sibling_63_segment_cells,
                               "pseudotimed_BTN_mother_sibling_73_segment_cells" = pseudotimed_BTN_mother_sibling_73_segment_cells)

modified_URD <- add_segment_state(BTN_segment_cells_list, "BTN_segment_cells_list")
plotTree(modified_URD, label.segments = F, label = "segment_state", label.type = "group")
```

# (2) Get PC Coordinates and Corresponding Pseudotimes for BTN Cells 
The PC coordinates and corresponding pseudotimes for BTN Cells will help determine
where to cut the segments and find pseudostates for BTN Precursors, their Sibling(s) and Mothers. 
Pseudostates are approximations of true Precursor, Precursor Sibling(s) and Mother states. They will
be used to calculate Differentially Expressed Genes in the next part of the pipeline.

## Make Seurat Object
```{r eval=FALSE}
BTN_cells <- c(pseudotimed_BTN_precursor_segment_cells, pseudotimed_BTN_precursor_sibling_segment_cells, 
               pseudotimed_BTN_mother_segment_cells, pseudotimed_BTN_mother_sibling_63_segment_cells, 
               pseudotimed_BTN_mother_sibling_73_segment_cells)
  
BTN_raw_mat <- URD_raw_data[, BTN_cells]
BTN_seurat <- CreateSeuratObject(counts = BTN_raw_mat,
                                min.cells = 3,
                                min.features = 200)
# Normalize Data.
BTN_seurat <- NormalizeData(BTN_seurat)
# Find Variable Feature.
BTN_seurat <- FindVariableFeatures(BTN_seurat, selection.method = "vst", nfeatures = 2000)
# Scale data.
all_BTN_seurat_genes <- rownames(BTN_seurat)
# use.umi argument to regress on umi count data.
BTN_seurat <- ScaleData(BTN_seurat, features = all_BTN_seurat_genes, use.umi = TRUE)
# Do PCA.
BTN_seurat <- RunPCA(BTN_seurat, features = VariableFeatures(object = BTN_seurat))
```

## Use Elbow Plot to choose the number of PCs (15)
```{r eval=FALSE}
ElbowPlot(BTN_seurat)
```

## Key Variables
```{r eval=FALSE}
BTN_PCA_COORDINATES <- Embeddings(BTN_seurat, reduction = "pca")
BTN_NUM_PCS <- 15
```

# (3) Get Pseudostates
As described, pseudostates are approximations of true Precursor, Precursor Sibling,
and Mother populations. We will first make "PC_pseudotime_dfs", which are data frames
that contain PC coordinates and pseudotimes for the BTN cells. 

The data frames of Precursor Precursor Sibling, and Mother segments will exported as 
csv files, which will then be used as inputs to modules in "get_pseudostates.py".

PC_pseudotime_dfs of Mother's Siblings will be subjected to a separate function (get_est_mother_sibling_cells)
to get their pseudostates.

## Make PC_pseudotime_dfs
```{r eval=FALSE}
BTN_precursor_segment_PC_pseudotime_df <- make_PC_pseudotime_df(pseudotimed_BTN_precursor_segment_cells, BTN_PCA_COORDINATES, BTN_NUM_PCS, pseudotimes)

BTN_sibling_segment_PC_pseudotime_df <- make_PC_pseudotime_df(pseudotimed_BTN_precursor_sibling_segment_cells, BTN_PCA_COORDINATES, BTN_NUM_PCS, pseudotimes)

BTN_mother_segment_PC_pseudotime_df <- make_PC_pseudotime_df(pseudotimed_BTN_mother_segment_cells, BTN_PCA_COORDINATES, BTN_NUM_PCS, pseudotimes)
```

## Export PC_pseudotime_dfs
```{r eval=FALSE}
PC_pseudotime_dfs_out_dir <- "./PC_pseudotime_dfs/"

write.table(BTN_precursor_segment_PC_pseudotime_df, file = paste0(PC_pseudotime_dfs_out_dir, "BTN_precursor_segment_PC_pseudotime_df.csv"), sep = "\t", row.names = FALSE)

write.table(BTN_sibling_segment_PC_pseudotime_df, file = paste0(PC_pseudotime_dfs_out_dir, "BTN_sibling_segment_PC_pseudotime_df.csv"), sep = "\t", row.names = FALSE)

write.table(BTN_mother_segment_PC_pseudotime_df, file = paste0(PC_pseudotime_dfs_out_dir, "BTN_mother_segment_PC_pseudotime_df.csv"), sep = "\t", row.names = FALSE)
```

## Load "get_pseudostates.py"
"get_pseudostates.py" consists of modules that, given a PC_pseudotime_dict of cells on a segment,
finds a cutoff that demarcates the pseudostate on that segment. That is, all the cells
on a segment until that cutoff are considered a part of the pseudostate.

Refer to Code documentation and Methods of paper for details on implementation details.

### Create a new reticulate environment 
This is where we will install all necessary python packages.
use
```{r eval=FALSE}
# create a new environment 
conda_create("r-reticulate")
use_condaenv("r-reticulate")
```

### Install Relevant Packages for "get_pseudostates.py" into given r-reticulate environment.
```{r eval=FALSE}
py_install("python-igraph", pip = T)
conda_install("r-reticulate", "scipy")
```

```{r eval=FALSE}
source_python("get_pseudostates.py")
```

## Key Variables
> n_leiden_iterations = number of iterations to be used by the Leiden algorithm.

>n_routine_iterations = number of iterations for the entire routine.

>verbose = refers to whether cutoffs will be printed as routine runs or not.

Please refer to Code documentation for further details.
```{r eval=FALSE}
n_leiden_iterations <- 5L
n_routine_iterations <- 50L
verbose <- FALSE
```

## BTN Precursor
```{r eval=FALSE}
set.seed(0)
# Took ~7 mins to run.
BTN_precursor_segment_df_file <- "./PC_pseudotime_dfs/BTN_precursor_segment_PC_pseudotime_df.csv"

BTN_precursor_segment_cutoff_output = get_pseudostates_pipeline(BTN_precursor_segment_df_file, n_leiden_iterations, n_routine_iterations, verbose)
```

```{r eval=FALSE}
set.seed(0)
BTN_precursor_pseudostate_cutoff <- BTN_precursor_segment_cutoff_output[[2]]
# 318
BTN_precursor_pseudostate_cells <- pseudotimed_BTN_precursor_segment_cells[1:BTN_precursor_pseudostate_cutoff]
```

## BTN Precursor Sibling
```{r eval=FALSE}
set.seed(0)
# Took ~11 mins to run.
BTN_precursor_sibing_segment_df_file <- "./PC_pseudotime_dfs/BTN_sibling_segment_PC_pseudotime_df.csv"
BTN_precursor_segment_cutoff_output <- get_pseudostates_pipeline(BTN_precursor_sibing_segment_df_file, n_leiden_iterations, n_routine_iterations, verbose)
```

```{r eval=FALSE}
set.seed(0)
BTN_precursor_sibling_pseudostate_cutoff <- BTN_precursor_segment_cutoff_output[[2]]
# 256
BTN_precursor_sibling_pseudostate_cells <- pseudotimed_BTN_precursor_sibling_segment_cells[1:BTN_precursor_sibling_pseudostate_cutoff]
```

## BTN Mother
```{r eval=FALSE}
set.seed(0)
# Took ~7 mins to run.
BTN_mother_segment_df_file <- "./PC_pseudotime_dfs/BTN_mother_segment_PC_pseudotime_df.csv"
BTN_mother_segment_cutoff_output <- get_pseudostates_pipeline(BTN_mother_segment_df_file, n_leiden_iterations, n_routine_iterations, verbose)
```

```{r eval=FALSE}
set.seed(0)
BTN_mother_pseudostate_cutoff <- BTN_mother_segment_cutoff_output[[2]]
# 240
BTN_mother_pseudostate_cells <- pseudotimed_BTN_mother_segment_cells[1:BTN_mother_pseudostate_cutoff]
```

## Mother's Siblings
The Mother's Siblings must be comparable to it in pseudotime. Thus, we use the 
“get_mother_sibling_pseudostates” function to find their pseudostates. 
We combine them into one pseudostate.

```{r eval=FALSE}
# 280
BTN_mother_sibling_63_output <- get_mother_sibling_pseudostates(BTN_mother_pseudostate_cells, pseudotimed_BTN_mother_sibling_63_segment_cells)

mother_sibling_63_first_index <- BTN_mother_sibling_63_output$first_index
mother_sibling_63_last_index <- BTN_mother_sibling_63_output$last_index

# 12
BTN_mother_sibling_73_output <- get_mother_sibling_pseudostates(BTN_mother_pseudostate_cells, pseudotimed_BTN_mother_sibling_73_segment_cells)

mother_sibling_73_first_index <- BTN_mother_sibling_73_output$first_index
mother_sibling_73_last_index <- BTN_mother_sibling_73_output$last_index

# 292
BTN_mother_sibling_pseudostate_cells <- c(pseudotimed_BTN_mother_sibling_63_segment_cells[mother_sibling_63_first_index:mother_sibling_63_last_index],              pseudotimed_BTN_mother_sibling_73_segment_cells[mother_sibling_73_first_index:mother_sibling_73_last_index])
```

## Visualize Pseudostates on Virtual Neural Lineage Tree
```{r eval=FALSE}
BTN_pseudostates_list <- list("BTN_precursor_pseudostate_cells" = BTN_precursor_pseudostate_cells,
                              "BTN_precursor_sibling_pseudostate_cells" = BTN_precursor_sibling_pseudostate_cells,
                              "BTN_mother_pseudostate_cells" = BTN_mother_pseudostate_cells,
                              "BTN_mother_sibling_pseudostate_cells" = BTN_mother_sibling_pseudostate_cells)

modified_URD <- add_segment_state(BTN_pseudostates_list, "BTN_pseudostates_list")
plotTree(modified_URD, label.segments = F, label = "segment_state", label.type = "group")
```

# (4) Get Differentially Expressed Genes
After identifying the Precursor, Mother, and their respective Sibling(s) populations, we obtain the following gene lists by calculating Differentially Expressed Genes between the specified populations.

### Precursor Identity Genes
>These are genes that define the overall state of the Precursor. They will be activated by the Cocktail TFs.

Comparison 1: Precursor vs. Sibling

### Precursor DEG TFs
>This set contains TFs relevant for specification that are present in the Precursor.

Comparison 2: Precursor vs. Sibling(s)

Comparison 3: Precursor vs. Mother

> The DEG TFs that result from the comparison between the Precursor and its Mother are those that get activated in the Precursor from the previous time point.

### Mother DEG TFs
>This set contains TFs relevant for specification that are present in the Mother.
Comparison 4: Mother vs. Sibling(s)

Finally, we will take the union of DEG TFs from Comparisons 2, 3, and 4 and refer  to it as "Unfiltered Candidate Cocktail TFs".

## Raw Matrices
```{r eval=FALSE}
# Raw Matrices with expression levels of only TFs. 
# 595 x 318
BTN_precursor_pseudostate_TF_raw_mat <- URD_raw_data[tfs_in_URD, BTN_precursor_pseudostate_cells]
# 595 x 256
BTN_precursor_sibling_pseudostate_TF_raw_mat <- URD_raw_data[tfs_in_URD, BTN_precursor_sibling_pseudostate_cells]
# 595 x 240
BTN_mother_pseudostate_TF_raw_mat <- URD_raw_data[tfs_in_URD, BTN_mother_pseudostate_cells]
# 595 x 292
BTN_mother_sibling_pseudostate_TF_raw_mat <- URD_raw_data[tfs_in_URD, BTN_mother_sibling_pseudostate_cells]

# Raw Matrices with expression levels of all genes.
# 14809 x 318
BTN_precursor_pseudostate_raw_mat <- URD_raw_data[, BTN_precursor_pseudostate_cells]
# 14809 x 256
BTN_precursor_sibling_pseudostate_raw_mat <- URD_raw_data[, BTN_precursor_sibling_pseudostate_cells]
```

All DEG lists will be written out to this dir.
```{r eval=FALSE}
DEG_lists_out_dir <- "./DEG_lists/"
```

## Precursor vs. Precursor Sibling- DEG TFs and Precursor Identity Genes
### Precursor Identity Genes
Perform edgeR to get DEGs between Precursor and Precursor Sibling Pseudostates.
```{r eval=FALSE}
BTN_precursor_v_sibling_pseudostate_mat <- as.matrix(cbind(BTN_precursor_pseudostate_raw_mat, BTN_precursor_sibling_pseudostate_raw_mat))
```

```{r eval=FALSE}
BTN_precursor_v_sibling_pseudostate_edgeR_outputs <- complete_edgeR_pipeline(BTN_precursor_v_sibling_pseudostate_mat, ncol(BTN_precursor_pseudostate_raw_mat), ncol(BTN_precursor_sibling_pseudostate_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 215
BTN_precursor_pseudostate_DEGs_df <- BTN_precursor_v_sibling_pseudostate_edgeR_outputs[[1]]
BTN_precursor_pseudostate_DEGs <- rownames(BTN_precursor_pseudostate_DEGs_df)
```

#### Write out Precursor DEGs (Precursor Identity Genes)
```{r eval=FALSE}
write_csv_helper(BTN_precursor_pseudostate_DEGs, paste0(DEG_lists_out_dir, "BTN_precursor_pseudostate_DEGs.csv"))
```

### Precursor DEG TFs
```{r eval=FALSE}
BTN_precursor_v_sibling_pseudostate_TF_mat <- as.matrix(cbind(BTN_precursor_pseudostate_TF_raw_mat, BTN_precursor_sibling_pseudostate_TF_raw_mat))

BTN_precursor_v_sibling_pseudostate_TF_edgeR_outputs <- complete_edgeR_pipeline(BTN_precursor_v_sibling_pseudostate_TF_mat, ncol(BTN_precursor_pseudostate_TF_raw_mat), ncol(BTN_precursor_sibling_pseudostate_TF_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 12
BTN_precursor_v_sibling_pseudostate_TF_DEGs_df <- BTN_precursor_v_sibling_pseudostate_TF_edgeR_outputs[[1]]
# Get TF orthologs.
named_BTN_precursor_v_sibling_pseudostate_TF_DEGs_df <- convert_khids_to_ortholog(BTN_precursor_v_sibling_pseudostate_TF_DEGs_df)

BTN_precursor_v_sibling_pseudostate_TF_DEGs <- named_BTN_precursor_v_sibling_pseudostate_TF_DEGs_df$Ghost.Name
```

#### Write out TF DEGs
```{r eval=FALSE}
write_csv_helper(BTN_precursor_v_sibling_pseudostate_TF_DEGs, paste0(DEG_lists_out_dir, "BTN_precursor_v_sibling_pseudostate_TF_DEGs.csv"))
```

## Precursor vs. Mother- DEG TFs
```{r eval=FALSE}
BTN_precursor_v_mother_pseudostate_TF_mat <- as.matrix(cbind(BTN_precursor_pseudostate_TF_raw_mat, BTN_mother_pseudostate_TF_raw_mat))

BTN_precursor_v_mother_pseudostate_TF_edgeR_outputs <- complete_edgeR_pipeline(BTN_precursor_v_mother_pseudostate_TF_mat, ncol(BTN_precursor_pseudostate_TF_raw_mat), ncol(BTN_mother_pseudostate_TF_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 13
BTN_precursor_v_mother_pseudostate_TF_DEGs_df <- BTN_precursor_v_mother_pseudostate_TF_edgeR_outputs[[1]]
# Get TF orthologs.
named_BTN_precursor_v_mother_pseudostate_TF_DEGs_df <- convert_khids_to_ortholog(BTN_precursor_v_mother_pseudostate_TF_DEGs_df)

BTN_precursor_v_mother_pseudostate_TF_DEGs <- named_BTN_precursor_v_mother_pseudostate_TF_DEGs_df$Ghost.Name
```

#### Write out TF DEGs
```{r eval=FALSE}
write_csv_helper(BTN_precursor_v_mother_pseudostate_TF_DEGs, paste0(DEG_lists_out_dir, "BTN_precursor_v_mother_pseudostate_TF_DEGs.csv"))
```

## Mother vs. Mother's Sibling- DEG TFs
```{r eval=FALSE}
BTN_mother_v_sibling_pseudostate_TF_mat <- as.matrix(cbind(BTN_mother_pseudostate_TF_raw_mat, BTN_mother_sibling_pseudostate_TF_raw_mat))

BTN_mother_v_sibling_pseudostate_TF_edgeR_outputs <- complete_edgeR_pipeline(BTN_mother_v_sibling_pseudostate_TF_mat, ncol(BTN_mother_pseudostate_TF_raw_mat), ncol(BTN_mother_sibling_pseudostate_TF_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 14
BTN_mother_v_sibling_pseudostate_TF_DEGs_df <- BTN_mother_v_sibling_pseudostate_TF_edgeR_outputs[[1]]
# Get TF orthologs.
named_BTN_mother_v_sibling_pseudostate_TF_DEGs_df <- convert_khids_to_ortholog(BTN_mother_v_sibling_pseudostate_TF_DEGs_df)

BTN_mother_v_sibling_pseudostate_TF_DEGs <- named_BTN_mother_v_sibling_pseudostate_TF_DEGs_df$Ghost.Name
```

#### Write out TF DEGs
```{r eval=FALSE}
write_csv_helper(BTN_mother_v_sibling_pseudostate_TF_DEGs, paste0(DEG_lists_out_dir, "BTN_mother_v_sibling_pseudostate_TF_DEGs.csv"))
```

# (5) Get Candidate Cocktail TFs
We assume that the most relevant Unfiltered Candidate Cocktail TFs in either the 
Precursor or Mother will be those that are most highly expressed. Furthermore,
based on the idea that TFs that are highly expressed in the Mother are likely to 
continue to be highly expressed in the Precursor, we will take the union of the top 50%
most highly expressed Unfiltered Candidate Cocktail TFs in either the Mother or
Precursor, and consider these the Candidate Cocktail TFs.

These Candidate Cocktail TFs, along with the Precursor Identity Genes will be
used as inputs to the Group Lasso Model. 

## Get Mean Expression of all DEG TFs
```{r eval=FALSE}
BTN_precursor_pseudostate_mean_TF_exp_df <- make_named_mean_TF_exp_df(BTN_precursor_pseudostate_cells)

BTN_mother_pseudostate_mean_TF_exp_df <- make_named_mean_TF_exp_df(BTN_mother_pseudostate_cells)
```

######## REF SUPP FIGURE PG 2 ##########
## Take the union of top 50% most highly expressed Unfiltered Candidate Cocktail TFs in Precursor and Mother. These are the Candidate Cocktail TFs
```{r eval=FALSE}
# 26
BTN_unfiltered_candidate_cocktail_tfs <- unique(c(named_BTN_precursor_v_sibling_pseudostate_TF_DEGs_df$KH.gene.model,
                                                  named_BTN_precursor_v_mother_pseudostate_TF_DEGs_df$KH.gene.model,
                                                  named_BTN_mother_v_sibling_pseudostate_TF_DEGs_df$KH.gene.model))

# Take 50% cutoff.
# This returns both a union and intersection of top 50% most highly expressed unfiltered candidate cocktail TFs.
BTN_cutoff_df_list <- get_cutoff_df(BTN_precursor_pseudostate_mean_TF_exp_df, BTN_mother_pseudostate_mean_TF_exp_df, BTN_unfiltered_candidate_cocktail_tfs, 0.5)

# Let us take the union.
# 15
BTN_cutoff_union <- BTN_cutoff_df_list$union
# These are the Candidate Cocktail TFs.
BTN_candidate_cocktail_tfs <- unique(c(BTN_cutoff_union$Ghost.Name_precursor, BTN_cutoff_union$Ghost.Name_mother))
# Remove any NA's!
BTN_candidate_cocktail_tfs <- BTN_candidate_cocktail_tfs[!is.na(BTN_candidate_cocktail_tfs)]

# We also want their khids.
BTN_candidate_cocktail_khids <- BTN_cutoff_union$KH.gene.model
```

### Write out the Candidate Cocktail TFs
```{r eval=FALSE}
write_csv_helper(BTN_candidate_cocktail_tfs, paste0(DEG_lists_out_dir, "BTN_candidate_cocktail_tfs.csv"))
```

# (6) Do Cocktail TF Inference using Group Lasso Model
The final component of the computational method is a Group Lasso Model in which
individual TFs are groups and subsets of these are selected at various values of
the regularization parameter lambda to "explain" expression of Precursor DEGs.
The model captures the idea that expression of Precursor DEGs will be regulated 
by a common set of TFs. We consider the first set of 3 TFs as the Cocktail.

Please refer to the Methods of paper for a full description of the model.

Briefly, Group Lasso Matrix X contains expression of Candidate Cocktail TFs in all Precursor cells; 
Matrix Y contains expression of Precursor DEGs in all Precursor cells;
The Group Lasso Model learns a coefficient vector that has "DEG-specific" TF coefficients, 
that is, it learns a unique set of TF weights for each Precursor DEG as each DEG
will be affected by the same TF to a different extent.

The Group Lasso Model is implemented using the grpreg package.

## Get a set of Precursor DEGs excluding DEG TFs
Their expression will be explained by the model.
```{r eval=FALSE}
# 4 
common_BTN_DEG_TFs <- intersect(BTN_candidate_cocktail_khids, BTN_precursor_pseudostate_DEGs)
# 211
cleaned_BTN_precursor_pseudostate_DEGs <- setdiff(BTN_precursor_pseudostate_DEGs, common_BTN_DEG_TFs)
```

## Get numbers of Precursor cells (n), Candidate Cocktail TFs (p), and Precusor DEGs (k) 
```{r eval=FALSE}
# 318, 15, 211
BTN_n <- length(BTN_precursor_pseudostate_cells)
BTN_p <- length(BTN_candidate_cocktail_tfs)
BTN_k <- length(cleaned_BTN_precursor_pseudostate_DEGs)
```

## Make Group Lasso Matrix X
Matrix S is a matrix that has the expression of all Candidate Cocktail TFs in all Precursor Cells of dimensions n x p (below). It will be repeated k times along the diagonal of Group Lasso Matrix X. Thus, Group Lasso Matrix X will have dimensions nk x pk. All other entries in Group Lasso Matrix X will be 0. 

### Make Matrix S 
Matrix S is the same as BTN_TFs_in_precursor_pseudostate_mat below.

```{r eval=FALSE}
# 318 x 15
BTN_TFs_in_precursor_pseudostate_mat <- t(URD_data[BTN_candidate_cocktail_khids, BTN_precursor_pseudostate_cells])
```

### Concatenate k S Matrices to make Group Lasso Matrix X
```{r eval=FALSE}
BTN_S_matrix <- BTN_TFs_in_precursor_pseudostate_mat
BTN_zero_matrix <- matrix(0, BTN_n, BTN_p)
# This is a dummy matrix- we will remove it at the end.
BTN_grplasso_X_mat <- matrix(1, BTN_n, BTN_p*BTN_k)

for (i in 1:BTN_k) {
  BTN_L_matrix <- do.call(cbind, replicate(i-1, BTN_zero_matrix, simplify=FALSE))
  BTN_R_matrix <- do.call(cbind, replicate(BTN_k-i, BTN_zero_matrix, simplify=FALSE))
  # At each iteration, we create a row in which Matrix S is sandwiched between 
  # Left "L" Zero Matrix(ces) and Right "R" Zero Matrix(ces) given its position on
  # the diagonal of Group Lasso Matrix X.
  BTN_grplasso_mat_row <- do.call(cbind, list(BTN_L_matrix, BTN_S_matrix, BTN_R_matrix))
  BTN_grplasso_X_mat <- rbind(BTN_grplasso_X_mat, BTN_grplasso_mat_row)
}

# 67098 x 3165
BTN_grplasso_X_mat_2 <- BTN_grplasso_X_mat[-c(1:BTN_n), ]
```

### Rename Row and Column names for Convenience
#### Rename Row Names to reflect cell number and its DEG status 
The Row Names have the following pattern: the first n rows are DEG_1_cell_1...
DEG_1_cell_n. The following n rows are DEG_2_cell_1... DEG_2_cell_n for all k DEGs.

```{r eval=FALSE}
BTN_grplasso_X_mat_row_cells <- rep(paste0("_cell_", seq(1, BTN_n)), BTN_k)
BTN_DEG_statuses <- rep(c(1:BTN_k), each = BTN_n)
BTN_new_grplasso_X_mat_2_rownames <- paste0("DEG_", BTN_DEG_statuses, BTN_grplasso_X_mat_row_cells)

rownames(BTN_grplasso_X_mat_2) <- BTN_new_grplasso_X_mat_2_rownames
```

#### Rename Column Names to reflect TF number and its DEG status
Similarly, the Column Names have the following pattern: the first p rows are DEG_1_TF1...
DEG_1_TFp. The following p rows are DEG_2_TF1... DEG_2_TFp for all p TFs.

```{r eval=FALSE}
BTN_grplasso_X_mat_col_TFs <- rep(paste0("_TF", seq(1, BTN_p)), BTN_k)
BTN_DEG_statuses_cols <- rep(c(1:BTN_k), each = BTN_p)
BTN_new_grplasso_X_mat_2_colnames <- paste0("DEG_", BTN_DEG_statuses_cols, BTN_grplasso_X_mat_col_TFs)

colnames(BTN_grplasso_X_mat_2) <- BTN_new_grplasso_X_mat_2_colnames
```

### Final Group Lasso Matrix X
```{r eval=FALSE}
# 67098 x 3165
final_grplasso_X_mat <- as.matrix(BTN_grplasso_X_mat_2)
```

## Make Group Lasso Matrix Y
Group Lasso Matrix Y contains the expression of all Precursor DEGs in all Precursor Cells in the following way: The first n rows contain expression of DEG 1 in all n cells; the next n rows contain expression of DEG 2 in all n cells; so and so forth for k DEGs.
It has dimensions nk x 1.

### Make a matrix that has the expression of all Precursor DEGs in Precursor Cells of dimensions k x n
```{r eval=FALSE}
# 318 x 211
BTN_precursor_DEGs_in_precursors_mat <- t(URD_data[cleaned_BTN_precursor_pseudostate_DEGs, BTN_precursor_pseudostate_cells])
```

### Index above matrix so as so make Group Lasso Matrix Y
```{r eval=FALSE}
BTN_grplasso_Y_vec <- c()

for (DEG in 1:BTN_k) {
  for (cell in 1:BTN_n) {
    BTN_DEG_cell_exp <- BTN_precursor_DEGs_in_precursors_mat[cell, DEG]
    BTN_grplasso_Y_vec <- c(BTN_grplasso_Y_vec, BTN_DEG_cell_exp)
  }
}

# 67098 x 1
BTN_grplasso_Y_mat <- as.matrix(BTN_grplasso_Y_vec, ncol=1, nrow=(BTN_n*BTN_k))
```

### Rename Row names to reflect cell and DEG status
The Row Names follow the same pattern as the Row Names of Group Lasso Matrix X.

```{r eval=FALSE}
BTN_grplasso_Y_DEG_statuses <- rep(c(1:BTN_k), each = BTN_n)
BTN_grplasso_Y_DEGs <- paste0("DEG_", BTN_grplasso_Y_DEG_statuses)
BTN_grplasso_Y_cells <- rep(paste0("_cell_", seq(1, BTN_n)), BTN_k)
BTN_new_grplasso_Y_rownames <- paste0(BTN_grplasso_Y_DEGs, BTN_grplasso_Y_cells)

rownames(BTN_grplasso_Y_mat) <- BTN_new_grplasso_Y_rownames
```

### Final Matrix Y
```{r eval=FALSE}
# 67098 x 1
final_BTN_grplasso_Y_mat <- BTN_grplasso_Y_mat
```

## Run Group Lasso
### Variables: Group Lasso Matrices X and Y and Group IDs
Each TF is a group. Thus, there are p groups each of size k. Given the way Group Lasso Matrix X is organized, the Group IDs follow the pattern of 1, 2, … p; 1, 2, … p; k times. 

```{r eval=FALSE}
grplasso_BTN_groups <- rep(seq(1, BTN_p), BTN_k)

BTN_X <- final_grplasso_X_mat
BTN_Y <- final_BTN_grplasso_Y_mat
BTN_group_id <- grplasso_BTN_groups
```

### Do Cross-Validation
```{r eval=FALSE}
# ~54 mins to Run 
grplasso_BTN_cv_grepreg_fit <- cv.grpreg(BTN_X, BTN_Y, BTN_group_id, penalty="grLasso")
```

######## REF SUPP FIGURE PG 3 ##########
### Get Sets of Selected TF Groups at Various Values of Lambda
```{r eval=FALSE}
# Key: lambda; Values: Set of Selected TF Groups
BTN_group_lassoed_tf_list <- get_group_lassoed_tf_list(grplasso_BTN_cv_grepreg_fit, BTN_candidate_cocktail_tfs)

# Key: lambda; Values: Set of Non-selected TF Groups
BTN_non_group_lassoed_TF_list <- lapply(BTN_group_lassoed_tf_list, get_non_group_lassoed_TFs, BTN_candidate_cocktail_tfs)
```

#### Let us Visualize the Number of Groups Selected at Various Values of Lambda
```{r eval=FALSE}
plot(grplasso_BTN_cv_grepreg_fit)
```

### Get Sorted Total TF Importance Lists
The Total Importance Score of a Cocktail TF is the sum of the absolute value of its DEG-specific coefficients. It thus gives a quantitative measure of a Cocktail TF's strength over the Gene Regulatory Network (as defined by the Precursor DEGs)
underlying the Precursor's state. 

```{r eval=FALSE}
# (For all lambdas) key = lambda; value is a list with key = TF and value = non0 betas.
BTN_TF_importances_lists <- get_TF_importances_list(grplasso_BTN_cv_grepreg_fit)

# Sum the absolute values of these betas for all TFs for all lambdas.
BTN_sum_total_TF_importance_lists <- BTN_TF_importances_lists %>% map(get_total_summed_TF_importance_list)

# Make sorted dfs given these Total Importance Scores.
BTN_sorted_sum_total_TF_importance_lists <- BTN_sum_total_TF_importance_lists %>% map(make_sorted_TF_importance_df, BTN_candidate_cocktail_tfs)
```

# Final Result: Inferred BTN Cocktail
Upon Inspection of "BTN_group_lassoed_tf_list", we see that the TF Set of size 3 consists of Neurogenin, COE, and HNF6 ("Collier BTN Cocktail").

TF Set of size 6 consists of these 3 TFs along with Orphan bHLH-1, msxb, and POU4 ("POU4 BTN Cocktail").

# Extended Analyses 
>Extended Analyses 1 and 2 compare Cocktail Inference results from another Linear Model, Gaussian Multivariate Multi-Response Linear Regression (GMML), and a Nonlinear Model based on GRNBoost2 respectively.

> Extended Analysis 3 compares Cocktail Inference results with Terminal BTNs (fully differentiated BTNs).

> Extended Analysis 4 performs subsampling of BTN Precursor data to build confidence for Group Lasso results. 

## (1) Comparison to Gaussian Multivariate Multi-Response Linear Regression (GMML)
DEG expression levels can be modeled as linear combinations of Candidate Cocktail TF expression levels. In our case, we have multiple predictors (TFs) and multiple response variables (expression of DEGs) and both the predictors and response variables are correlated. Thus, an appropriate way to model these interactions is via a gaussian multivariate multi-response linear regression (GMML). 

To infer the Cocktail, we employ 2 strategies:  

Strategy 1: We calculate the average coefficient values for each Candidate
Cocktail TF, get a Total Importance Score (described afterwards), and consider the top 3 based on
the Total Importance Score as the Cocktail.   

Strategy 2: We repeat the same process, but on the **ranks** of the coefficient
values for each Candidate Cocktail TF. 
This is because DEGs with very high expression levels will bias coefficient values
to be higher. We prefer Strategy 2.   

Refer to Methods of paper for a full description of the model.

### Matrices X and Y
Matrix X has expression of Cocktail Candidate TFs in Precursor cells and 
has dimensions n x p. 
Matrix Y has expression of all Precursor DEGs in Precursor cells and
has dimensions n x k.

```{r eval=FALSE}
glm_BTN_X_mat <- as.matrix(BTN_TFs_in_precursor_pseudostate_mat)
glm_BTN_Y_mat <- as.matrix(BTN_precursor_DEGs_in_precursors_mat)
```

### Perform Regression
#### Without Lasso penalty
We will refer to this as the complete GMML model.
```{r eval=FALSE}
# ~Took 15s to run.
BTN_fit <- cv.glmnet(glm_BTN_X_mat, glm_BTN_Y_mat, intercept = FALSE, family = "mgaussian")
BTN_lambda <- BTN_fit$lambda.min
# BTN_coef is a list with keys = khids of Precursor DEGs and values = series of attributes
# one of which (@ x) contains weights for all 15 input TF variables.
BTN_coef_list <- coef(BTN_fit, s = BTN_lambda)
```

#### With Lasso penalty, selecting for 3 TF Variables
Here, we set the dfmax parameter to 2, which applies a Lasso penalty and selects for 3 TF variables.

One can observe the results given different values of this parameter, such as 3 or 4,
which respectively returns 4 or 5 variables, but we study 3 variables in order to 
compare this result to our inferred cocktail, which contains 3 TFs.

```{r eval=FALSE}
# Also took ~15s to run.
BTN_fit_3TFs <- cv.glmnet(glm_BTN_X_mat, glm_BTN_Y_mat, intercept = FALSE, family = "mgaussian", dfmax = 2)
BTN_lambda_3TFs <- BTN_fit_3TFs$lambda.min
BTN_coef_3TFs_list <- coef(BTN_fit_3TFs, s = BTN_lambda_3TFs)
```

### Strategy 1
#### Get Regression Coefficients for all Candidate Cocktail TFs
Make a coef_df, which is a data frame with rows = Candidate Cocktail Precursor DEGs, 
columns = Precursor DEGs, and entries as regression coefficients present in coef_list.

```{r eval=FALSE}
BTN_coef_df <- make_coef_df(BTN_coef_list,  BTN_candidate_cocktail_tfs, cleaned_BTN_precursor_pseudostate_DEGs)

BTN_coef_df_3TFs <- make_coef_df(BTN_coef_3TFs_list,  BTN_candidate_cocktail_tfs, cleaned_BTN_precursor_pseudostate_DEGs)
```

#### Visualize Distribution of Coefficients and get their Averages
We notice that HNF6, COE, and Pou4 have the highest average coefficient values in the complete GMML model. 

######## REF SUPP FIGURE PG 3, 5 ##########
```{r eval=FALSE}
BTN_average_coef_df <- get_average_coefficient_df(BTN_coef_df)
BTN_coef_distribution_plot <- get_coefficients_distribution_plot(BTN_coef_df)
BTN_coef_distribution_plot

BTN_average_coef_df_3TFs <- get_average_coefficient_df(BTN_coef_df_3TFs)
```

#### Get Total Importance Scores for Candidate Cocktail TFs
> Sum the absolute values of the regression coefficients

######## REF SUPP FIGURE PG 7 ##########
```{r eval=FALSE}
BTN_total_GMML_importance_df <- get_total_GMML_importance_df(BTN_coef_df)
BTN_total_GMML_importance_df_3TFs <- get_total_GMML_importance_df(BTN_coef_df_3TFs)
```

### Strategy 2
#### First Get the ranks of the absolute values of the regression coefficients

```{r eval=FALSE}
ranked_BTN_coef_df <- BTN_coef_df %>% mutate(across(everything(), abs), across(everything(), min_rank))
ranked_BTN_coef_df_3TFs <- BTN_coef_df_3TFs %>% mutate(across(everything(), abs), across(everything(), min_rank))
```

#### Visualize Distribution of Coefficients and get their Averages
We notice that HNF6, COE, and POU4 have the highest ranked average coefficient values in the complete GMML model. 

######## REF SUPP FIGURE PG 4, 6 ##########
```{r eval=FALSE}
ranked_BTN_average_coef_df <- get_average_coefficient_df(ranked_BTN_coef_df)
ranked_BTN_coef_distribution_plot <- get_coefficients_distribution_plot(ranked_BTN_coef_df)
ranked_BTN_coef_distribution_plot

ranked_BTN_average_coef_df_3TFs <- get_average_coefficient_df(ranked_BTN_coef_df_3TFs)
```

#### Get Total Importance Scores for Candidate Cocktail TFs

######## REF SUPP FIGURE PG 8 ##########
```{r eval=FALSE}
ranked_BTN_total_GMML_importance_df <- get_total_GMML_importance_df(ranked_BTN_coef_df)
ranked_BTN_total_GMML_importance_df_3TFs <- get_total_GMML_importance_df(ranked_BTN_coef_df_3TFs)
```

### Notes on Final Result
> As Neurogenin is missing, the BTN Cocktail is not correctly inferred from the complete 
GMML model. This demonstrates the need for the Group Lasso.

## (2) Comparison with GRNBoost2 Inferences
The purpose of this part of the analysis is to compare Cocktail predictions derived
from nonlinear models (GRNBoost2) with those from the linear Group Lasso model, 
as TF-gene relationships may be better modeled as nonlinear ones.

This Pipeline returns 2 lists:
sorted_total_importance_tfs and sorted_percent_top_list, which are used in tandem
to infer the Cocktail. Please refer to documentation in "GRNBoost2_pipeline.py"
to see the methods behind deriving the lists.

```{r eval=FALSE}
BTN_GRN_dir <- "GRN/BTN/"
```

#### BTN Precursor DEG Counts Matrix
```{r eval=FALSE}
# Matrix must have counts from both DEGs and Candidate Cocktail TFs.
# 226
BTN_precursor_pseudostate_DEGs_plus <- unique(c(BTN_precursor_pseudostate_DEGs, BTN_candidate_cocktail_khids))
# 226 x 318
BTN_precursor_pseudostate_GRN_mat <- URD_raw_data[BTN_precursor_pseudostate_DEGs_plus, BTN_precursor_pseudostate_cells]
# 318 x 226
BTN_precursor_pseudostate_GRN_df <- as.data.frame(t(BTN_precursor_pseudostate_GRN_mat))
```

#### Inputs to GRNBoost2 Pipeline
```{r eval=FALSE}
write_tsv(BTN_precursor_pseudostate_GRN_df, paste0("GRN/", "BTN_precursor_pseudostate_GRN_df.tsv"))
```

```{r eval=FALSE}
write_csv_helper(BTN_candidate_cocktail_khids, paste0("GRN/", "BTN_candidate_cocktail_khids.csv"))
```

#### Do GRNBoost2 50 times.
### Install Relevant Packages for "GRNBoost2_pipeline.py" into given r-reticulate environment.
```{r eval=FALSE}
conda_install("r-reticulate", "pandas")
conda_install("r-reticulate", "numpy")
py_install("arboreto", pip = T)
```

```{r eval=FALSE}
source_python("GRNBoost2_pipeline.py")
```

```{r eval=FALSE}
set.seed(16)

expression_count_matrix_tsv <- paste0("GRN/", "BTN_precursor_pseudostate_GRN_df.tsv")
tf_khid_csv <- paste0("GRN/", "BTN_candidate_cocktail_khids.csv")
outname <- "GRN/BTN/BTN"
min_i <- 0L
max_i <- 50L

# ~Took 12 mins to run.
do_GRNBoost2(expression_count_matrix_tsv, tf_khid_csv, outname, min_i, max_i)
```

##### sorted_total_importance_tfs
Each element of the list is a tuple of TF and list with Mean and Standard Deviation
of Total Importance Scores. The TFs are sorted in nonincreasing order based on the mean.
           
##### sorted_percent_top_list
Each element of the list is a TF and its Percentage Covered as a Top 3 Regulator for all DEGs.

######## REF SUPP FIGURE PG 8, 9 ##########
```{r eval=FALSE}
khid_tf_tsv_file <- "./helper_files/Ciona_khid_TF.tsv"
BTN_GRN_path <- "/Users/prakritipaul/Git/Cocktails/FINAL_code_and_results/GRN/BTN/*.tsv"
n <- 3L

# Returns sorted_total_importance_tfs and sorted_percent_top_list
BTN_GRNBoost2_outputs <- GRNBoost2_pipeline(BTN_GRN_path, khid_tf_tsv_file, n)
```

##### Inspect Output contents.
```{r eval=FALSE}
BTN_sorted_total_importance_tfs <- BTN_GRNBoost2_outputs[1]
BTN_sorted_percent_top_list <- BTN_GRNBoost2_outputs[2]
```

### Notes on Final Result
> The Cocktail appears within the first 4 TFs, which supports the results obtained from the Group Lasso Model.

## (3) Comparison with Method Applied on Terminal BTNs
A necessary comparison is with Terminal BTNs, or fully differentiated BTNs, as 
a key claim in this work is that the Cocktail cannot be inferred without 
studying cell states at the time of specification. 

Potentially, we can find the Cocktail either in the top 50% mostly highly expressed DEG TFs in the Terminal DEGs (these are Terminal Candidate Cocktail TFs) or in the output of running the Group Lasso Model using Terminal DEGs and Terminal Candidate Cocktail Factors.
We will also use the GMML.

From the following analyses, we demonstrate that the Cocktail cannot be inferred from either way and build evidence for the power of our computational method and the strength of the assumptions we made.

### Identify Terminal BTNs and their Siblings
Siblings were identified from observing the Virtual Lineage Tree.

```{r eval=FALSE}
# 65
terminal_BTNs <- neural_CT_list$BTNs
terminal_BTN_siblings <- c(neural_CT_list$CENs, neural_CT_list$RTENs, neural_CT_list$aATENs)
```

### Raw Mats
```{r eval=FALSE}
# 14809 x 65
terminal_BTN_raw_mat <- URD_raw_data[, terminal_BTNs]
# 595 x 65
terminal_BTN_TF_raw_mat <- URD_raw_data[tfs_in_URD, terminal_BTNs]
# 14809 x 286
terminal_BTN_sibling_raw_mat <- URD_raw_data[, terminal_BTN_siblings]
# 595 x 286
terminal_BTN_sibling_TF_raw_mat <- URD_raw_data[tfs_in_URD, terminal_BTN_siblings]
```

### Calculate Terminal DEGs with respect to Siblings
Terminal DEGs are analogous to Precursor DEGs. 

We calculate Terminal DEG TFs with respect to Siblings because comparison to all other neural
cell types did not render meaningful TFs. Instead, TFs known to play roles in specification appeared as differentially expressed when Terminal BTNs were compared to its siblings. This made these results more competitive with the results obtained from applying the computational method to Precursors at the time of specification.

### Get Terminal DEGs
```{r eval=FALSE}
terminal_BTN_v_siblings_mat <- as.matrix(cbind(terminal_BTN_raw_mat, terminal_BTN_sibling_raw_mat))

terminal_BTN_v_siblings_edgeR_outputs <- complete_edgeR_pipeline(terminal_BTN_v_siblings_mat, ncol(terminal_BTN_raw_mat), ncol(terminal_BTN_sibling_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 713
terminal_BTN_v_siblings_DEGs_df <- terminal_BTN_v_siblings_edgeR_outputs[[1]]
terminal_BTN_v_siblings_DEGs <- rownames(terminal_BTN_v_siblings_DEGs_df)
```

#### Write them out.
```{r eval=FALSE}
write_csv_helper(terminal_BTN_v_siblings_DEGs, paste0(DEG_lists_out_dir, "terminal_BTN_v_siblings_DEGs.csv"))
```

### Get Terminal DEG TFs
We will filter these downstream to get Terminal Candidate Cocktail TFs.
```{r eval=FALSE}
terminal_BTN_v_siblings_TF_mat <- as.matrix(cbind(terminal_BTN_TF_raw_mat, terminal_BTN_sibling_TF_raw_mat))

terminal_BTN_v_siblings_TF_edgeR_outputs <- complete_edgeR_pipeline(terminal_BTN_v_siblings_TF_mat, ncol(terminal_BTN_TF_raw_mat), ncol(terminal_BTN_sibling_TF_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 22
terminal_BTN_v_siblings_TF_DEGs_df <- terminal_BTN_v_siblings_TF_edgeR_outputs[[1]]
# Get TF orthologs.
named_terminal_BTN_v_siblings_TF_DEGs_df <- convert_khids_to_ortholog(terminal_BTN_v_siblings_TF_DEGs_df)
terminal_BTN_v_sibling_TF_DEGs <- named_terminal_BTN_v_siblings_TF_DEGs_df$Ghost.Name
```

#### Write them out.
```{r eval=FALSE}
write_csv_helper(terminal_BTN_v_sibling_TF_DEGs, paste0(DEG_lists_out_dir, "terminal_BTN_v_sibling_TF_DEGs.csv"))
```

### Get Terminal Candidate Cocktail TFs
As before, we will take the top 50% most highly expressed of the Terminal DEG TFs and consider them as the Terminal Candidate Cocktail TFs.

######## REF SUPP FIGURE PG 10 ##########
```{r eval=FALSE}
# Get mean Expression of all DEG TFs
terminal_BTN_mean_TF_exp_df <- make_named_mean_TF_exp_df(terminal_BTNs)

# Get khids of Terminal DEG TFs
terminal_BTN_v_siblings_TF_DEG_khids <- rownames(terminal_BTN_v_siblings_TF_DEGs_df)

# Get Mean expression of Terminal DEG TFs in terminal cells.
terminal_BTN_cocktail_TF_mean_exp_df <- terminal_BTN_mean_TF_exp_df %>% filter(KH.gene.model %in% terminal_BTN_v_siblings_TF_DEG_khids)

# Get top 50% most highly expressed of above DEG TFS
terminal_cutoff <- 0.5

terminal_BTN_cutoff_df <- terminal_BTN_cocktail_TF_mean_exp_df[1:(terminal_cutoff*(length(terminal_BTN_v_siblings_TF_DEG_khids))), ]

# Get the TFs and khids
terminal_BTN_candidate_cocktail_khids <- terminal_BTN_cutoff_df$KH.gene.model
terminal_BTN_candidate_cocktail_tfs <- terminal_BTN_cutoff_df$Ghost.Name
```

#### Write out the Candidate Cocktail TFs
```{r eval=FALSE}
write_csv_helper(terminal_BTN_candidate_cocktail_tfs, paste0(DEG_lists_out_dir, "terminal_BTN_candidate_cocktail_tfs.csv"))
```

#### Brief Aside
An interesting observation is that the common Candidate Cocktail TFs between the Precursor and Terminal BTN are the Cocktail itself!
```{r eval=FALSE}
common_BTN_cocktail_TFs <- intersect(terminal_BTN_candidate_cocktail_tfs, BTN_candidate_cocktail_tfs)
common_BTN_cocktail_TFs
```

### (a) Do Cocktail TF Inference using Group Lasso Model
#### Variables
```{r eval=FALSE}
# Get rid of TFs in DEGs.
# 11
common_terminal_BTN_DEG_TFs <- intersect(terminal_BTN_candidate_cocktail_khids, terminal_BTN_v_siblings_DEGs)
# 702
cleaned_terminal_BTN_DEGs <- setdiff(terminal_BTN_v_siblings_DEGs, common_terminal_BTN_DEG_TFs)
```

```{r eval=FALSE}
# 65, 11, 702
terminal_BTN_n <- length(terminal_BTNs)
terminal_BTN_p <- length(terminal_BTN_candidate_cocktail_tfs)
terminal_BTN_k <- length(cleaned_terminal_BTN_DEGs)
```

#### Make Group Lasso X Matrix
```{r eval=FALSE}
# 65 x 11
terminal_BTN_TFs_in_terminal_mat <- t(URD_data[terminal_BTN_candidate_cocktail_khids, terminal_BTNs])
```

```{r eval=FALSE}
# 45630 x 7722
terminal_BTN_grplasso_X_mat <- make_grplasso_X_mat(terminal_BTN_TFs_in_terminal_mat, terminal_BTN_n, terminal_BTN_p, terminal_BTN_k)
```

#### Make Group Lasso Y Matrix
```{r eval=FALSE}
# 65 x 702
terminal_BTN_DEGs_in_terminal_mat <- t(URD_data[cleaned_terminal_BTN_DEGs, terminal_BTNs])
```

```{r eval=FALSE}
# 45630 x 1
terminal_BTN_grplasso_Y_mat <- make_grplasso_Y_mat(terminal_BTN_DEGs_in_terminal_mat, terminal_BTNs, terminal_BTN_p, terminal_BTN_k)
```

#### Variables: Group Lasso Matrices X and Y and Group IDs
```{r eval=FALSE}
terminal_grplasso_BTN_groups <- rep(seq(1, terminal_BTN_p), terminal_BTN_k)

terminal_BTN_X <- terminal_BTN_grplasso_X_mat
terminal_BTN_Y <- terminal_BTN_grplasso_Y_mat
terminal_BTN_group_id <- terminal_grplasso_BTN_groups
```

#### Run Group Lasso 
```{r eval=FALSE}
# Took ~4h to run!
terminal_grplasso_BTN_cv_grepreg_fit <- cv.grpreg(terminal_BTN_X, terminal_BTN_Y, terminal_BTN_group_id, penalty="grLasso")
```

### Get Sets of Selected TF Groups at Various Values of Lambda
While Neurogenin appears in the first set of 5, other Cocktail TFs (POU4, HNF6, and COE) do not. Thus, Cocktail cannot be uncovered.

######## REF SUPP FIGURE PG 11 ##########
```{r eval=FALSE}
# Key: lambda; Values: Set of Selected TF Groups
terminal_BTN_group_lassoed_tf_list <- get_group_lassoed_tf_list(terminal_grplasso_BTN_cv_grepreg_fit, terminal_BTN_candidate_cocktail_tfs)

# Key: lambda; Values: Set of Non-selected TF Groups
terminal_BTN_non_group_lassoed_TF_list <- lapply(terminal_BTN_group_lassoed_tf_list, get_non_group_lassoed_TFs, terminal_BTN_candidate_cocktail_tfs)
```

#### Let us Visualize the Number of Groups Selected at Various Values of Lambda
```{r eval=FALSE}
plot(terminal_grplasso_BTN_cv_grepreg_fit)
```

### Get Sorted Total TF Importance Lists
The Total Importance Scores within these lists give us a quantitative measure of a 
Cocktail TF's strength over the Gene Regulatory Network (as defined by the Precursor DEGs)
underlying the Precursor's state. 

```{r eval=FALSE}
# (For all lambdas) key = lambda; value is a list with key = TF_num and value = betas.
terminal_BTN_TF_importances_lists <- get_TF_importances_list(terminal_grplasso_BTN_cv_grepreg_fit)

# Sum the absolute values of these betas for all TFs for all lambdas.
terminal_BTN_sum_total_TF_importance_lists <- terminal_BTN_TF_importances_lists %>% map(get_total_summed_TF_importance_list)

# Make sorted dfs given these Total Importance Scores.
terminal_BTN_sorted_sum_total_TF_importance_lists <- terminal_BTN_sum_total_TF_importance_lists %>% map(make_sorted_TF_importance_df, terminal_BTN_candidate_cocktail_tfs)
```

#### Result
We are unable to uncover either BTN Cocktail. 

### (b) Do GMML 
```{r eval=FALSE}
glm_terminal_BTN_X_mat <- as.matrix(terminal_BTN_TFs_in_terminal_mat)
glm_terminal_BTN_Y_mat <- as.matrix(terminal_BTN_DEGs_in_terminal_mat)
```

### Perform Regression
We will refer to this as the complete GMML model.
```{r eval=FALSE}
# ~Took 40s to run.
terminal_BTN_fit <- cv.glmnet(glm_terminal_BTN_X_mat, glm_terminal_BTN_Y_mat, intercept = FALSE, family = "mgaussian")
terminal_BTN_lambda <- terminal_BTN_fit$lambda.min
# terminal_BTN_coef is a list with keys = khids of Precursor DEGs and values = series of attributes
# one of which (@ x) contains weights for all 14 input TF variables.
terminal_BTN_coef_list <- coef(terminal_BTN_fit, s = terminal_BTN_lambda)
```

### Strategy 1
#### Get Regression Coefficients for all Candidate Cocktail TFs
Make a coef_df, which is a data frame with rows = Candidate Cocktail Precursor DEGs, 
columns = Precursor DEGs, and entries as regression coefficients present in coef_list.

```{r eval=FALSE}
terminal_BTN_coef_df <- make_coef_df(terminal_BTN_coef_list,  terminal_BTN_candidate_cocktail_tfs, cleaned_terminal_BTN_DEGs)

terminal_BTN_coef_df_3TFs <- make_coef_df(terminal_BTN_coef_3TFs_list,  terminal_BTN_candidate_cocktail_tfs, cleaned_terminal_BTN_DEGs)
```

#### Visualize Distribution of Coefficients and get their Averages
We notice that PUR-a/b/y, FLJ21616-related, and Ci-ZF169 have the highest ranked average coefficient values in the complete GMML model. 

######## REF SUPP FIGURE PG 12, 13 ##########
```{r eval=FALSE}
terminal_BTN_average_coef_df <- get_average_coefficient_df(terminal_BTN_coef_df)
terminal_BTN_coef_distribution_plot <- get_coefficients_distribution_plot(terminal_BTN_coef_df)

terminal_BTN_average_coef_df_3TFs <- get_average_coefficient_df(terminal_BTN_coef_df_3TFs)
```

#### Get Total Importance Scores for Candidate Cocktail TFs
######## REF SUPP FIGURE PG 14 ##########
```{r eval=FALSE}
terminal_terminal_BTN_total_GMML_importance_df <- get_total_GMML_importance_df(terminal_BTN_coef_df)

terminal_BTN_total_GMML_importance_df_3TFs <- get_total_GMML_importance_df(terminal_BTN_coef_df_3TFs)
```

### Strategy 2
#### First Get the ranks of the absolute values of the regression coefficients
######## REF SUPP FIGURE PG 12, 13 ##########
```{r eval=FALSE}
ranked_terminal_BTN_coef_df <- terminal_BTN_coef_df %>% mutate(across(everything(), abs), across(everything(), min_rank))

ranked_terminal_BTN_coef_df_3TFs <- terminal_BTN_coef_df_3TFs %>% mutate(across(everything(), abs), across(everything(), min_rank))
```

#### Visualize Distribution of Coefficients and get their Averages
We notice that PUR-a/b/y, FLJ21616-related, and Ci-ZF169 have the highest ranked average coefficient values in the complete GMML model. 

```{r eval=FALSE}
ranked_terminal_BTN_average_coef_df <- get_average_coefficient_df(ranked_terminal_BTN_coef_df)
ranked_terminal_BTN_coef_distribution_plot <- get_coefficients_distribution_plot(ranked_terminal_BTN_coef_df)
ranked_terminal_BTN_coef_distribution_plot

ranked_terminal_BTN_average_coef_df_3TFs <- get_average_coefficient_df(ranked_terminal_BTN_coef_df_3TFs)
```

#### Get Total Importance Scores for Candidate Cocktail TFs

######## REF SUPP FIGURE PG 14 ##########
```{r eval=FALSE}
ranked_terminal_BTN_total_GMML_importance_df <- get_total_GMML_importance_df(ranked_terminal_BTN_coef_df)
ranked_terminal_BTN_total_GMML_importance_df_3TFs <- get_total_GMML_importance_df(ranked_terminal_BTN_coef_df_3TFs)
```

### (c) Comparison with GRNBoost2 Inferences
The purpose of this part of the analysis is to compare Cocktail predictions derived
from nonlinear models (GRNBoost2) with those from the linear Group Lasso model, 
as TF-gene relationships may be better modeled as nonlinear ones.

This Pipeline returns 2 lists:
sorted_total_importance_tfs and sorted_percent_top_list, which are used in tandem
to infer the Cocktail. Please refer to documentation in "GRNBoost2_pipeline.py"
to see the methods behind deriving the lists.

```{r eval=FALSE}
terminal_BTN_GRN_dir <- "./GRN/terminal_BTN/"
```

##### Terminal BTN DEG Counts Matrix
```{r eval=FALSE}
# Matrix must have counts from both DEGs and Candidate Cocktail TFs.
terminal_BTN_DEGs_plus <- unique(c(terminal_BTN_v_siblings_DEGs, terminal_BTN_candidate_cocktail_khids))
terminal_BTN_GRN_mat <- URD_raw_data[terminal_BTN_DEGs_plus, terminal_BTNs]

# 65 x 713
terminal_BTN_GRN_df <- as.data.frame(t(terminal_BTN_GRN_mat))
```

#### Inputs to GRNBoost2 Pipeline
```{r eval=FALSE}
write_tsv(terminal_BTN_GRN_df, paste0("GRN/", "terminal_BTN_GRN_df.tsv"))
```

```{r eval=FALSE}
write_csv_helper(terminal_BTN_candidate_cocktail_khids, paste0("GRN/", "terminal_BTN_candidate_cocktail_khids.csv"))
```

#### Do GRNBoost2 50 times.
```{r eval=FALSE}
set.seed(16)

expression_count_matrix_tsv <- paste0("GRN/", "terminal_BTN_GRN_df.tsv")
tf_khid_csv <- paste0("GRN/", "terminal_BTN_candidate_cocktail_khids.csv")
outname <- "GRN/terminal_BTN/terminal_BTN"
min_i <- 0L
max_i <- 50L

# Took ~16 mins to run.
do_GRNBoost2(expression_count_matrix_tsv, tf_khid_csv, outname, min_i, max_i)
```

##### sorted_total_importance_tfs
Each element of the list is a tuple of TF and list with Mean and Standard Deviation
of Total Importance Scores. The TFs are sorted in nonincreasing order based on the mean.
           
##### sorted_percent_top_list
Each element of the list is a TF and its Percentage Covered as a Top 6 Regulator for all DEGs.

######## REF SUPP FIGURE PG 15 ##########
```{r eval=FALSE}
khid_tf_tsv_file <- "./helper_files/Ciona_khid_TF.tsv"
terminal_BTN_GRN_path <- "./GRN/terminal_BTN/*.tsv"
n <- 3L

# Returns sorted_total_importance_tfs and sorted_percent_top_list
terminal_BTN_GRNBoost2_outputs <- GRNBoost2_pipeline(terminal_BTN_GRN_path, khid_tf_tsv_file, n)
```

##### Inspect Output contents.
```{r eval=FALSE}
terminal_BTN_sorted_total_importance_tfs <- terminal_BTN_GRNBoost2_outputs[1]
terminal_BTN_sorted_percent_top_list <- terminal_BTN_GRNBoost2_outputs[2]
```

#### Note on Final Result
As our Cocktail TFs are not among the top TFs given these two measures (only Neurogenin does), we are once again able to uncover the Cocktail from studying the terminal state. This further supports the need to study gene regulatory changes occurring at the time of specification.

## (4) Subsampling
> Finally, in order to build evidence for our predictions, we randomly subsample 90%
of the data, run the Group Lasso Model, and keep track of its inferred Cocktail/TF Set of an ideal size (Materials and Methods). We count the number of times that a Cocktail TF  appears in a subsample's inferred Cocktail/TF Set and thus get a percentage or confidence for the overall prediction. 

### Get 50 runs of Group Lasso Model on Subsampled Data.
> The results of this subsampling can be found in the given directory- each 
object has 5 runs of cv.grpreg, resulting in 50 samples/runs. 
The following code makes a “combined_subsampled_tf_df”, a data frame with
row = run_*_lambda_*,  where run = the round of subsampling/run of group lasso
and lambda = one of 100 values of lambda used during cross-validation; 
and column value = string that is a concatenated name of the set of TF Groups
selected at the corresponding value of lambda ("TF-string").
These "combined_subsampled_tf_df"s are subsequently processed in  "subsampling_prediction_confidence.py".

##### Access subsamples.
```{r include=TRUE}
if (!(file.exists("/tmp/BTN_subsamples"))) {
  cat("Downloading BTN_subsamples directory\n")
  options(timeout = 600)
  BTN_subsampling_link <- "https://cocktails-r-objects.s3.amazonaws.com/BTN_subsamples.zip"
  # Download in /tmp.
  download.file(BTN_subsampling_link, destfile = "/tmp/BTN_subsamples.zip")
  cat("Unzipping\n")
  unzip("/tmp/BTN_subsamples.zip", exdir = "/tmp")
  
} else {
  cat("BTN_subsamples directory is present.\n")
}
```
##### Make combined_subsampled_tf_df.
```{r eval=FALSE}
BTN_combined_subsampled_tf_df <- make_combined_subsampled_tf_df("/tmp/BTN_subsamples/")
```

```{r eval=FALSE}
subsampling_out_dir <- "./subsampling/"
```

##### Write out combined_subsampled_tf_df.
```{r eval=FALSE}
write_out_combined_subsampled_tf_df(BTN_combined_subsampled_tf_df, subsampling_out_dir, "BTN_combined_subsampled_tf_df.csv")
```

##### Also write out BTN's group_lassoed_tf_list (Has Predicted Cocktail).
```{r eval=FALSE}
write_out_CT_group_lassoed_tf_list(BTN_group_lassoed_tf_list, subsampling_out_dir, "BTN_group_lassoed_tf_df.csv")
```

### Run Subsampling Prediction Confidence Pipeline.
```{r eval=FALSE}
source_python("subsampling_prediction_confidence.py")
```

```{r eval=FALSE}
BTN_group_lassoed_tf_df_csv <- paste0(subsampling_out_dir, "BTN_group_lassoed_tf_df.csv")
BTN_subsampled_tf_df_csv <- paste0(subsampling_out_dir, "BTN_combined_subsampled_tf_df.csv")
BTN_tf_set_sizes <- list(3, 4, 5, 6)
BTN_ideal_set_size <- 6

BTN_cocktail_subsampling_confidence <- subsampling_prediction_confidence(BTN_group_lassoed_tf_df_csv, BTN_subsampled_tf_df_csv, BTN_tf_set_sizes, BTN_ideal_set_size)
```

### Note on Final Result
Neurogenin, HNF6, and COE each appear 100% time in TF Sets of size 3.  
Neurogenin, HNF6, COE, and POU4 each appear 100% time in TF Sets of size 6.

# RSession Info
> Refer to Supplement for complete information.

```{r include=TRUE}
sessionInfo()
```

