---
title: "GLRA Cocktail Inference Pipeline"

subtitle: "This RMarkdown includes all analyses resulting from the computational methods developed in <insert name of Paper and authors> to infer the ddN Cocktail"

author: "Written by Prakriti Paul Chacha"
  
output:
  github_document:
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Import Relevant Libraries
```{r eval=FALSE, error=FALSE, warning=FALSE}
library(URD)
library(Seurat)
library(ggplot2)
library(tidyverse)
library(stats)
library(Matrix)
library(stringr)
library(edgeR)
library(rlang)
library(grpreg)
library(glmnet)
library(reticulate)
source("./helper_files/final_main_pipeline_start_script.R")
source("./helper_files/final_helper_functions.R")
```

# (1) Identify Precursor, Mother, and their Sibling(s) Segment Cell Populations on Virtual Neural Lineage Tree
Segment cells refers to the cells along the tree segment/branch that will contain
Precursor, Precursor Sibling, and Mother pseudostates.

Note: "CNS.k300s6w4.tree.built.name" is the URD object generated from time-course
      single-cell RNA-seq data in Chen et. al* for the Central Nervous System (CNS)
      of Ciona intestinalis.
      It contains cell names, their pseudotimes, the Virtual Lineage Tree, and all
      URD object attributes.

*Cao, Chen, et al. "Comprehensive single-cell transcriptome lineages of a proto-vertebrate." Nature 571.7765 (2019): 349-354.

> Note: GLRA refers to the cell segment that contains both GLRA and non-GLRA cells.
GLRAs express Dmbx+, while the latter does not. Thus, throughout this pipeline,
GLRA ID's and siblings will be used, but later subsetted to get GLRAs.

## Precursor Segment 

### Get IDs of GLRA Precursor Segment and its Sibling Segment
```{r eval=FALSE}
# Get ID from URD Object.
GLRA_ID <- "38"
# 31, 45
GLRA_sibling_IDs <- segSiblings(CNS.k300s6w4.tree.built.name, GLRA_ID, include.self = F)
GLRA_sibling_31_ID <- GLRA_sibling_IDs[1]
GLRA_sibling_45_ID <- GLRA_sibling_IDs[2]
```

### Get GLRA Precursor and Sibling Segment Cells
>Key Note: GLRA_segment_cells contain both ddN (Dmbx+ cells) and non-ddN cells. We will need to differentiate between the two below in order to correctly identify the dd Precursor Pseudostate.

```{r eval=FALSE}
# 82
GLRA_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", GLRA_ID)
# 139
GLRA_sibling_31_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", GLRA_sibling_31_ID)
# 27 (Will need to take everything.) 
GLRA_sibling_45_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", GLRA_sibling_45_ID)
```

### Order Cells by Pseudotime
```{r eval=FALSE}
pseudotimed_GLRA_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, GLRA_ID, length(GLRA_segment_cells), ascending=TRUE)

# Get GLRA cells that express Dmbx- these are the GLRAs.
# 16/82
pseudotimed_ddN_precursor_cells <- get_dmbx_expressing_cells(pseudotimed_GLRA_segment_cells, pseudotimes)

pseudotimed_GLRA_sibling_31_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, GLRA_sibling_31_ID, length(GLRA_sibling_31_segment_cells), ascending=TRUE)

pseudotimed_GLRA_sibling_45_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, GLRA_sibling_45_ID, length(GLRA_sibling_45_segment_cells), ascending=TRUE)
```

## Mother Segment

### Get IDs of GLRA Mother Segment and its Siblings
```{r eval=FALSE}
GLRA_mother_ID <- "59"
# 64, 15
GLRA_mother_sibling_IDs <- segSiblings(CNS.k300s6w4.tree.built.name, GLRA_mother_ID, include.self = F)
GLRA_mother_sibling_64_ID <- GLRA_mother_sibling_IDs[1]
GLRA_mother_sibling_15_ID <- GLRA_mother_sibling_IDs[2]
```

### Get GLRA Mother and Sibling Segment Cells
```{r eval=FALSE}
# 395
GLRA_mother_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", GLRA_mother_ID)
# 147
GLRA_mother_sibling_64_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", GLRA_mother_sibling_64_ID)
# 89
GLRA_mother_sibling_15_segment_cells <- cellsInCluster(CNS.k300s6w4.tree.built.name, clustering="segment", GLRA_mother_sibling_15_ID)
```

### Order Cells by Pseudotime
```{r eval=FALSE}
pseudotimed_GLRA_mother_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, GLRA_mother_ID, length(GLRA_mother_segment_cells), ascending=FALSE)

pseudotimed_GLRA_mother_sibling_64_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, GLRA_mother_sibling_64_ID, length(GLRA_mother_sibling_64_segment_cells), ascending=FALSE)

pseudotimed_GLRA_mother_sibling_15_segment_cells <- get_pseudotimed_cells(cells_in_segments_list, pseudotimes, GLRA_mother_sibling_15_ID, length(GLRA_mother_sibling_15_segment_cells), ascending=FALSE)
```

## Visualize Cells on Virtual Neural Lineage Tree
```{r eval=FALSE}
GLRA_segment_cells_list <- list("pseudotimed_ddN_precursor_cells" = pseudotimed_ddN_precursor_cells,
                               "pseudotimed_GLRA_sibling_31_segment_cells" = pseudotimed_GLRA_sibling_31_segment_cells,
                               "pseudotimed_GLRA_sibling_45_segment_cells" = pseudotimed_GLRA_sibling_45_segment_cells,
                               "pseudotimed_GLRA_mother_segment_cells" = pseudotimed_GLRA_mother_segment_cells,
                               "pseudotimed_GLRA_mother_sibling_64_segment_cells" = pseudotimed_GLRA_mother_sibling_64_segment_cells,
                               "pseudotimed_GLRA_mother_sibling_15_segment_cells" = pseudotimed_GLRA_mother_sibling_15_segment_cells)

modified_URD <- add_segment_state(GLRA_segment_cells_list, "GLRA_segment_cells_list")
plotTree(modified_URD, label.segments = F, label = "segment_state", label.type = "group")
```

# (2) Get PC Coordinates and Corresponding Pseudotimes for GLRA Cells 
The PC coordinates and corresponding pseudotimes for GLRA Cells will help determine
where to cut the segments and find pseudostates for ddN Precursors, GLRA Sibling(s) and Mothers. 
Pseudostates are approximations of true Precursor, Precursor Sibling(s) and Mother states. They will
be used to calculate Differentially Expressed Genes in the next part of the pipeline.

## Make Seurat Object
```{r eval=FALSE}
GLRA_cells <- c(pseudotimed_ddN_precursor_cells, 
                pseudotimed_GLRA_sibling_31_segment_cells,
                pseudotimed_GLRA_sibling_45_segment_cells,
                pseudotimed_GLRA_mother_segment_cells, 
                pseudotimed_GLRA_mother_sibling_64_segment_cells, 
                pseudotimed_GLRA_mother_sibling_15_segment_cells)
  
GLRA_raw_mat <- URD_raw_data[, GLRA_cells]
GLRA_seurat <- CreateSeuratObject(counts = GLRA_raw_mat,
                                min.cells = 3,
                                min.features = 200)
# Normalize Data.
GLRA_seurat <- NormalizeData(GLRA_seurat)
# Find Variable Feature.
GLRA_seurat <- FindVariableFeatures(GLRA_seurat, selection.method = "vst", nfeatures = 2000)
# Scale data.
all_GLRA_seurat_genes <- rownames(GLRA_seurat)
# use.umi argument to regress on umi count data.
GLRA_seurat <- ScaleData(GLRA_seurat, features = all_GLRA_seurat_genes, use.umi = TRUE)
# Do PCA.
GLRA_seurat <- RunPCA(GLRA_seurat, features = VariableFeatures(object = GLRA_seurat))
```

## Use Elbow Plot to choose the number of PCs (15)
```{r eval=FALSE}
ElbowPlot(GLRA_seurat)
```

## Key Variables
```{r eval=FALSE}
GLRA_PCA_COORDINATES <- Embeddings(GLRA_seurat, reduction = "pca")
GLRA_NUM_PCS <- 15
```

# (3) Get Pseudostates
As described, pseudostates are approximations of true Precursor, Precursor Sibling,
and Mother populations. We will first make "PC_pseudotime_dfs", which are data frames
that contain PC coordinates and pseudotimes for the GLRA cells. 

The data frames of Precursor Precursor Sibling, and Mother segments will exported as 
csv files, which will then be used as inputs to modules in "get_pseudostates.py".

PC_pseudotime_dfs of Mother's Siblings will be subjected to a separate function (get_est_mother_sibling_cells)
to get their pseudostates.

## Make PC_pseudotime_dfs
```{r eval=FALSE}
ddN_PC_pseudotime_df <- make_PC_pseudotime_df(pseudotimed_ddN_precursor_cells, GLRA_PCA_COORDINATES, GLRA_NUM_PCS, pseudotimes)

GLRA_sibling_31_segment_PC_pseudotime_df <- make_PC_pseudotime_df(pseudotimed_GLRA_sibling_31_segment_cells, GLRA_PCA_COORDINATES, GLRA_NUM_PCS, pseudotimes)

GLRA_sibling_45_segment_PC_pseudotime_df <- make_PC_pseudotime_df(pseudotimed_GLRA_sibling_45_segment_cells, GLRA_PCA_COORDINATES, GLRA_NUM_PCS, pseudotimes)

GLRA_mother_segment_PC_pseudotime_df <- make_PC_pseudotime_df(pseudotimed_GLRA_mother_segment_cells, GLRA_PCA_COORDINATES, GLRA_NUM_PCS, pseudotimes)
```

## Export PC_pseudotime_dfs
```{r eval=FALSE}
PC_pseudotime_dfs_out_dir <- "./PC_pseudotime_dfs/"

write.table(ddN_PC_pseudotime_df, file = paste0(PC_pseudotime_dfs_out_dir, "ddN_PC_pseudotime_df.csv"), sep = "\t", row.names = FALSE)

write.table(GLRA_sibling_31_segment_PC_pseudotime_df, file = paste0(PC_pseudotime_dfs_out_dir, "GLRA_sibling_31_segment_PC_pseudotime_df.csv"), sep = "\t", row.names = FALSE)

write.table(GLRA_sibling_45_segment_PC_pseudotime_df, file = paste0(PC_pseudotime_dfs_out_dir, "GLRA_sibling_45_segment_PC_pseudotime_df.csv"), sep = "\t", row.names = FALSE)

write.table(GLRA_mother_segment_PC_pseudotime_df, file = paste0(PC_pseudotime_dfs_out_dir, "GLRA_mother_segment_PC_pseudotime_df.csv"), sep = "\t", row.names = FALSE)
```

## Load "get_pseudostates.py"
"get_pseudostates.py" consists of modules that, given a PC_pseudotime_dict of cells on a segment,
finds a cutoff that demarcates the pseudostate on that segment. That is, all the cells
on a segment until that cutoff are considered a part of the pseudostate.

Refer to Code documentation and Methods of paper for details on implementation details.

```{r eval=FALSE}
source_python("get_pseudostates.py")
```

## Key Variables
> n_leiden_iterations = number of iterations to be used by the Leiden algorithm.

>n_routine_iterations = number of iterations for the entire routine.

>verbose = refers to whether cutoffs will be printed as routine runs or not.

Please refer to Code documentation for further details.

```{r eval=FALSE}
n_leiden_iterations <- 5L
n_routine_iterations <- 50L
verbose <- FALSE
```

## ddN Precursor 
```{r eval=FALSE}
set.seed(0)
# Took ~3s to run.
ddN_df_file <- "./PC_pseudotime_dfs/ddN_PC_pseudotime_df.csv"

ddN_precursor_segment_cutoff_output = get_pseudostates_pipeline(ddN_df_file, n_leiden_iterations, n_routine_iterations, verbose)
```

```{r eval=FALSE}
set.seed(0)
ddN_precursor_pseudostate_cutoff <- ddN_precursor_segment_cutoff_output[[2]]
# 9
ddN_precursor_pseudostate_cells <- pseudotimed_ddN_precursor_cells[1:ddN_precursor_pseudostate_cutoff]
```

## GLRA Precursor Sibling 31
```{r eval=FALSE}
set.seed(0)
# Took ~1.5 mins to run.
GLRA_precursor_sibling_31_segment_df_file <- "./PC_pseudotime_dfs/GLRA_sibling_31_segment_PC_pseudotime_df.csv"
GLRA_precursor_sibling_31_segment_cutoff_output <- get_pseudostates_pipeline(GLRA_precursor_sibling_31_segment_df_file, n_leiden_iterations, n_routine_iterations, verbose)
```

```{r eval=FALSE}
set.seed(0)
GLRA_precursor_sibling_31_pseudostate_cutoff <- GLRA_precursor_sibling_31_segment_cutoff_output[[2]]
# 80
GLRA_precursor_sibling_31_pseudostate_cells <- pseudotimed_GLRA_sibling_31_segment_cells[1:GLRA_precursor_sibling_31_pseudostate_cutoff]
```

## GLRA Precursor Sibling 45
```{r eval=FALSE}
set.seed(0)
# Took ~5s to run.
GLRA_precursor_sibling_45_segment_df_file <- "./PC_pseudotime_dfs/GLRA_sibling_45_segment_PC_pseudotime_df.csv"
GLRA_precursor_sibling_45_segment_cutoff_output <- get_pseudostates_pipeline(GLRA_precursor_sibling_45_segment_df_file, n_leiden_iterations, n_routine_iterations, verbose)
```

```{r eval=FALSE}
set.seed(0)
GLRA_precursor_sibling_45_pseudostate_cutoff <- GLRA_precursor_sibling_45_segment_cutoff_output[[2]]
# 17
GLRA_precursor_sibling_45_pseudostate_cells <- pseudotimed_GLRA_sibling_45_segment_cells[1:GLRA_precursor_sibling_45_pseudostate_cutoff]
```

```{r eval=FALSE}
# 80/139 + 17/27 = 97/166
GLRA_precursor_sibling_pseudostate_cells <- c(GLRA_precursor_sibling_31_pseudostate_cells, GLRA_precursor_sibling_45_pseudostate_cells)
```

## GLRA Mother
```{r eval=FALSE}
set.seed(0)
# Took ~4.5 mins to run.
GLRA_mother_segment_df_file <- "./PC_pseudotime_dfs/GLRA_mother_segment_PC_pseudotime_df.csv"
GLRA_mother_segment_cutoff_output <- get_pseudostates_pipeline(GLRA_mother_segment_df_file, n_leiden_iterations, n_routine_iterations, verbose)
```

```{r eval=FALSE}
set.seed(0)
GLRA_mother_pseudostate_cutoff <- GLRA_mother_segment_cutoff_output[[2]]
# 171
GLRA_mother_pseudostate_cells <- pseudotimed_GLRA_mother_segment_cells[1:GLRA_mother_pseudostate_cutoff]
```

## Mother's Siblings
The Mother's Siblings must be comparable to it in pseudotime. Thus, we use the 
“get_mother_sibling_pseudostates” function to find their pseudostates. 
We combine them into one pseudostate.

```{r eval=FALSE}
# 12
GLRA_mother_sibling_64_output <- get_mother_sibling_pseudostates(GLRA_mother_pseudostate_cells, pseudotimed_GLRA_mother_sibling_64_segment_cells)

mother_sibling_64_first_index <- GLRA_mother_sibling_64_output$first_index
mother_sibling_64_last_index <- GLRA_mother_sibling_64_output$last_index

# 15
GLRA_mother_sibling_15_output <- get_mother_sibling_pseudostates(GLRA_mother_pseudostate_cells, pseudotimed_GLRA_mother_sibling_15_segment_cells)

mother_sibling_15_first_index <- GLRA_mother_sibling_15_output$first_index
mother_sibling_15_last_index <- GLRA_mother_sibling_15_output$last_index

# 27
GLRA_mother_sibling_pseudostate_cells <- c(pseudotimed_GLRA_mother_sibling_64_segment_cells[mother_sibling_64_first_index:mother_sibling_64_last_index],              pseudotimed_GLRA_mother_sibling_15_segment_cells[mother_sibling_15_first_index:mother_sibling_15_last_index])

```

## Visualize Pseudostates on Virtual Neural Lineage Tree
```{r eval=FALSE}
GLRA_pseudostates_list <- list("ddN_precursor_pseudostate_cells" = ddN_precursor_pseudostate_cells,
                              "GLRA_precursor_sibling_pseudostate_cells" = GLRA_precursor_sibling_pseudostate_cells,
                              "GLRA_mother_pseudostate_cells" = GLRA_mother_pseudostate_cells,
                              "GLRA_mother_sibling_pseudostate_cells" = GLRA_mother_sibling_pseudostate_cells)

modified_URD <- add_segment_state(GLRA_pseudostates_list, "GLRA_pseudostates_list")
plotTree(modified_URD, label.segments = F, label = "segment_state", label.type = "group")
```

# (4) Get Differentially Expressed Genes
After identifying the Precursor, Mother, and their respective Sibling(s) populations, we obtain the following gene lists by calculating Differentially Expressed Genes between the specified populations.

### Precursor Identity Genes
>These are genes that define the overall state of the Precursor. They will be activated by the Cocktail TFs.

Comparison 1: Precursor vs. Sibling

### Precursor DEG TFs
>This set contains TFs relevant for specification that are present in the Precursor.

Comparison 2: Precursor vs. Sibling(s)

Comparison 3: Precursor vs. Mother

> The DEG TFs that result from the comparison between the Precursor and its Mother are those that get activated in the Precursor from the previous time point.

### Mother DEG TFs
>This set contains TFs relevant for specification that are present in the Mother.
Comparison 4: Mother vs. Sibling(s)

Finally, we will take the union of DEG TFs from Comparisons 2, 3, and 4 and refer  to it as "Unfiltered Candidate Cocktail TFs".

## Raw Matrices
```{r eval=FALSE}
# Raw Matrices with expression levels of only TFs. 
# 595 x 9
ddN_precursor_pseudostate_TF_raw_mat <- URD_raw_data[tfs_in_URD, ddN_precursor_pseudostate_cells]
# 97
GLRA_precursor_sibling_pseudostate_TF_raw_mat <- URD_raw_data[tfs_in_URD, GLRA_precursor_sibling_pseudostate_cells]
# 171
GLRA_mother_pseudostate_TF_raw_mat <- URD_raw_data[tfs_in_URD, GLRA_mother_pseudostate_cells]
# 27
GLRA_mother_sibling_pseudostate_TF_raw_mat <- URD_raw_data[tfs_in_URD, GLRA_mother_sibling_pseudostate_cells]

# Raw Matrices with expression levels of all genes.
# 14809 x 9
ddN_precursor_pseudostate_raw_mat <- URD_raw_data[, ddN_precursor_pseudostate_cells]
# 97
GLRA_precursor_sibling_pseudostate_raw_mat <- URD_raw_data[, GLRA_precursor_sibling_pseudostate_cells]
```

All DEG lists will be written out to this dir.
```{r eval=FALSE}
DEG_lists_out_dir <- "./DEG_lists/"
```

## Precursor vs. Precursor Sibling- DEG TFs and Precursor Identity Genes
### Precursor Identity Genes
Perform edgeR to get DEGs between Precursor and Precursor Sibling Pseudostates.
```{r eval=FALSE}
ddN_precursor_v_sibling_pseudostate_mat <- as.matrix(cbind(ddN_precursor_pseudostate_raw_mat, GLRA_precursor_sibling_pseudostate_raw_mat))
```

```{r eval=FALSE}
ddN_precursor_v_sibling_pseudostate_edgeR_outputs <- complete_edgeR_pipeline(ddN_precursor_v_sibling_pseudostate_mat, ncol(ddN_precursor_pseudostate_raw_mat), ncol(GLRA_precursor_sibling_pseudostate_raw_mat), 0.05, -1, 50)
```

Get the DEGs.
```{r eval=FALSE}
# 196
ddN_precursor_pseudostate_DEGs_df <- ddN_precursor_v_sibling_pseudostate_edgeR_outputs[[1]]
ddN_precursor_pseudostate_DEGs <- rownames(ddN_precursor_pseudostate_DEGs_df)
```

#### Write out Precursor DEGs (Precursor Identity Genes)
```{r eval=FALSE}
write_csv_helper(ddN_precursor_pseudostate_DEGs, paste0(DEG_lists_out_dir, "ddN_precursor_pseudostate_DEGs.csv"))
```

### Precursor DEG TFs
```{r eval=FALSE}
ddN_precursor_v_sibling_pseudostate_TF_mat <- as.matrix(cbind(ddN_precursor_pseudostate_TF_raw_mat, GLRA_precursor_sibling_pseudostate_TF_raw_mat))

ddN_precursor_v_sibling_pseudostate_TF_edgeR_outputs <- complete_edgeR_pipeline(ddN_precursor_v_sibling_pseudostate_TF_mat, ncol(ddN_precursor_pseudostate_TF_raw_mat), ncol(GLRA_precursor_sibling_pseudostate_TF_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 16
ddN_precursor_v_sibling_pseudostate_TF_DEGs_df <- ddN_precursor_v_sibling_pseudostate_TF_edgeR_outputs[[1]]
# Get TF orthologs.
named_ddN_precursor_v_sibling_pseudostate_TF_DEGs_df <- convert_khids_to_ortholog(ddN_precursor_v_sibling_pseudostate_TF_DEGs_df)

ddN_precursor_v_sibling_pseudostate_TF_DEGs <- named_ddN_precursor_v_sibling_pseudostate_TF_DEGs_df$Ghost.Name
```

#### Write out TF DEGs
```{r eval=FALSE}
write_csv_helper(ddN_precursor_v_sibling_pseudostate_TF_DEGs, paste0(DEG_lists_out_dir, "ddN_precursor_v_sibling_pseudostate_TF_DEGs.csv"))
```

## Precursor vs. Mother- DEG TFs
```{r eval=FALSE}
ddN_precursor_v_mother_pseudostate_TF_mat <- as.matrix(cbind(ddN_precursor_pseudostate_TF_raw_mat, GLRA_mother_pseudostate_TF_raw_mat))

ddN_precursor_v_mother_pseudostate_TF_edgeR_outputs <- complete_edgeR_pipeline(ddN_precursor_v_mother_pseudostate_TF_mat, ncol(ddN_precursor_pseudostate_TF_raw_mat), ncol(GLRA_mother_pseudostate_TF_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 12
ddN_precursor_v_mother_pseudostate_TF_DEGs_df <- ddN_precursor_v_mother_pseudostate_TF_edgeR_outputs[[1]]
# Get TF orthologs.
named_ddN_precursor_v_mother_pseudostate_TF_DEGs_df <- convert_khids_to_ortholog(ddN_precursor_v_mother_pseudostate_TF_DEGs_df)

ddN_precursor_v_mother_pseudostate_TF_DEGs <- named_ddN_precursor_v_mother_pseudostate_TF_DEGs_df$Ghost.Name
```

#### Write out TF DEGs
```{r eval=FALSE}
write_csv_helper(ddN_precursor_v_mother_pseudostate_TF_DEGs, paste0(DEG_lists_out_dir, "ddN_precursor_v_mother_pseudostate_TF_DEGs.csv"))
```

## Mother vs. Mother's Sibling- DEG TFs
```{r eval=FALSE}
GLRA_mother_v_sibling_pseudostate_TF_mat <- as.matrix(cbind(GLRA_mother_pseudostate_TF_raw_mat, GLRA_mother_sibling_pseudostate_TF_raw_mat))

GLRA_mother_v_sibling_pseudostate_TF_edgeR_outputs <- complete_edgeR_pipeline(GLRA_mother_v_sibling_pseudostate_TF_mat, ncol(GLRA_mother_pseudostate_TF_raw_mat), ncol(GLRA_mother_sibling_pseudostate_TF_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 17
GLRA_mother_v_sibling_pseudostate_TF_DEGs_df <- GLRA_mother_v_sibling_pseudostate_TF_edgeR_outputs[[1]]
# Get TF orthologs.
named_GLRA_mother_v_sibling_pseudostate_TF_DEGs_df <- convert_khids_to_ortholog(GLRA_mother_v_sibling_pseudostate_TF_DEGs_df)

GLRA_mother_v_sibling_pseudostate_TF_DEGs <- named_GLRA_mother_v_sibling_pseudostate_TF_DEGs_df$Ghost.Name
```

#### Write out TF DEGs
```{r eval=FALSE}
write_csv_helper(GLRA_mother_v_sibling_pseudostate_TF_DEGs, paste0(DEG_lists_out_dir, "GLRA_mother_v_sibling_pseudostate_TF_DEGs.csv"))
```

# (5) Get Candidate Cocktail TFs
We assume that the most relevant Unfiltered Candidate Cocktail TFs in either the 
Precursor or Mother will be those that are most highly expressed. Furthermore,
based on the idea that TFs that are highly expressed in the Mother are likely to 
continue to be highly expressed in the Precursor, we will take the union of the top 50%
most highly expressed Unfiltered Candidate Cocktail TFs in either the Mother or
Precursor, and consider these the Candidate Cocktail TFs.

These Candidate Cocktail TFs, along with the Precursor Identity Genes will be
used as inputs to the Group Lasso Model. 

## Get Mean Expression of all DEG TFs
```{r eval=FALSE}
ddN_precursor_pseudostate_mean_TF_exp_df <- make_named_mean_TF_exp_df(ddN_precursor_pseudostate_cells)

GLRA_mother_pseudostate_mean_TF_exp_df <- make_named_mean_TF_exp_df(GLRA_mother_pseudostate_cells)
```

## Take the union of top 50% most highly expressed Unfiltered Candidate Cocktail TFs in Precursor and Mother. These are the Candidate Cocktail TFs

######## REF SUPP FIGURE PG 30 ##########
```{r eval=FALSE}
# 33
ddN_unfiltered_candidate_cocktail_tfs <- unique(c(named_ddN_precursor_v_sibling_pseudostate_TF_DEGs_df$KH.gene.model,
                                                  named_ddN_precursor_v_mother_pseudostate_TF_DEGs_df$KH.gene.model,
                                                  named_GLRA_mother_v_sibling_pseudostate_TF_DEGs_df$KH.gene.model))

# Take 50% cutoff.
# This returns both a union and intersection of top 50% most highly expressed unfiltered candidate cocktail TFs.
ddN_cutoff_df_list <- get_cutoff_df(ddN_precursor_pseudostate_mean_TF_exp_df, GLRA_mother_pseudostate_mean_TF_exp_df, ddN_unfiltered_candidate_cocktail_tfs, 0.5)

# Let us take the union.
# 25
ddN_cutoff_union <- ddN_cutoff_df_list$union
# These are the Candidate Cocktail TFs.
ddN_candidate_cocktail_tfs <- unique(c(ddN_cutoff_union$Ghost.Name_precursor, ddN_cutoff_union$Ghost.Name_mother))
# Remove any NA's!
ddN_candidate_cocktail_tfs <- ddN_candidate_cocktail_tfs[!is.na(ddN_candidate_cocktail_tfs)]

# We also want their khids.
ddN_candidate_cocktail_khids <- ddN_cutoff_union$KH.gene.model
```

### Write out the Candidate Cocktail TFs
```{r eval=FALSE}
write_csv_helper(ddN_candidate_cocktail_tfs, paste0(DEG_lists_out_dir, "ddN_candidate_cocktail_tfs.csv"))
```

# (6) Do Cocktail TF Inference using Group Lasso Model
The final component of the computational method is a Group Lasso Model in which
individual TFs are groups and subsets of these are selected at various values of
the regularization parameter lambda to "explain" expression of Precursor DEGs.
The model captures the idea that expression of Precursor DEGs will be regulated 
by a common set of TFs. We consider the first set of 3 TFs as the Cocktail.

Please refer to the Methods of paper for a full description of the model.

Briefly, Group Lasso Matrix X contains expression of Candidate Cocktail TFs in all Precursor cells; 
Matrix Y contains expression of Precursor DEGs in all Precursor cells;
The Group Lasso Model learns a coefficient vector that has "DEG-specific" TF coefficients, 
that is, it learns a unique set of TF weights for each Precursor DEG as each DEG
will be affected by the same TF to a different extent.

The Group Lasso Model is implemented using the grpreg package.

## Get a set of Precursor DEGs excluding DEG TFs
Their expression will be explained by the model.
```{r eval=FALSE}
# 11
common_ddN_DEG_TFs <- intersect(ddN_candidate_cocktail_khids, ddN_precursor_pseudostate_DEGs)
# 185
cleaned_ddN_precursor_pseudostate_DEGs <- setdiff(ddN_precursor_pseudostate_DEGs, common_ddN_DEG_TFs)
```

## Get numbers of Precursor cells (n), Candidate Cocktail TFs (p), and Precusor DEGs (k) 
```{r eval=FALSE}
# 9, 25, 185
ddN_n <- length(ddN_precursor_pseudostate_cells)
ddN_p <- length(ddN_candidate_cocktail_tfs)
ddN_k <- length(cleaned_ddN_precursor_pseudostate_DEGs)
```

### Make Matrix S 
Matrix S is the same as ddN_TFs_in_precursor_pseudostate_mat below.

```{r eval=FALSE}
# 9 x 25
ddN_TFs_in_precursor_pseudostate_mat <- t(URD_data[ddN_candidate_cocktail_khids, ddN_precursor_pseudostate_cells])
```

### Concatenate k S Matrices to make Group Lasso Matrix X
```{r eval=FALSE}
ddN_S_matrix <- ddN_TFs_in_precursor_pseudostate_mat
ddN_zero_matrix <- matrix(0, ddN_n, ddN_p)
# This is a dummy matrix- we will remove it at the end.
ddN_grplasso_X_mat <- matrix(1, ddN_n, ddN_p*ddN_k)

for (i in 1:ddN_k) {
  ddN_L_matrix <- do.call(cbind, replicate(i-1, ddN_zero_matrix, simplify=FALSE))
  ddN_R_matrix <- do.call(cbind, replicate(ddN_k-i, ddN_zero_matrix, simplify=FALSE))
  # At each iteration, we create a row in which Matrix S is sandwiched between 
  # Left "L" Zero Matrix(ces) and Right "R" Zero Matrix(ces) given its position on
  # the diagonal of Group Lasso Matrix X.
  ddN_grplasso_mat_row <- do.call(cbind, list(ddN_L_matrix, ddN_S_matrix, ddN_R_matrix))
  ddN_grplasso_X_mat <- rbind(ddN_grplasso_X_mat, ddN_grplasso_mat_row)
}

# 1665 x 4625
ddN_grplasso_X_mat_2 <- ddN_grplasso_X_mat[-c(1:ddN_n), ]
```

### Rename Row and Column names for Convenience
#### Rename Row Names to reflect cell number and its DEG status 
The Row Names have the following pattern: the first n rows are DEG_1_cell_1...
DEG_1_cell_n. The following n rows are DEG_2_cell_1... DEG_2_cell_n for all k DEGs.

```{r eval=FALSE}
ddN_grplasso_X_mat_row_cells <- rep(paste0("_cell_", seq(1, ddN_n)), ddN_k)
ddN_DEG_statuses <- rep(c(1:ddN_k), each = ddN_n)
ddN_new_grplasso_X_mat_2_rownames <- paste0("DEG_", ddN_DEG_statuses, ddN_grplasso_X_mat_row_cells)

rownames(ddN_grplasso_X_mat_2) <- ddN_new_grplasso_X_mat_2_rownames
```

#### Rename Column Names to reflect TF number and its DEG status
Similarly, the Column Names have the following pattern: the first p rows are DEG_1_TF1...
DEG_1_TFp. The following p rows are DEG_2_TF1... DEG_2_TFp for all p TFs.

```{r eval=FALSE}
ddN_grplasso_X_mat_col_TFs <- rep(paste0("_TF", seq(1, ddN_p)), ddN_k)
ddN_DEG_statuses_cols <- rep(c(1:ddN_k), each = ddN_p)
ddN_new_grplasso_X_mat_2_colnames <- paste0("DEG_", ddN_DEG_statuses_cols, ddN_grplasso_X_mat_col_TFs)

colnames(ddN_grplasso_X_mat_2) <- ddN_new_grplasso_X_mat_2_colnames
```

### Final Group Lasso Matrix X
```{r eval=FALSE}
# 1665 x 4625
final_grplasso_X_mat <- as.matrix(ddN_grplasso_X_mat_2)
```

## Make Group Lasso Matrix Y
Group Lasso Matrix Y contains the expression of all Precursor DEGs in all Precursor Cells in the following way: The first n rows contain expression of DEG 1 in all n cells; the next n rows contain expression of DEG 2 in all n cells; so and so forth for k DEGs.
It has dimensions nk x 1.

### Make a matrix that has the expression of all Precursor DEGs in Precursor Cells of dimensions k x n
```{r eval=FALSE}
# 9 x 185
ddN_precursor_DEGs_in_precursors_mat <- t(URD_data[cleaned_ddN_precursor_pseudostate_DEGs, ddN_precursor_pseudostate_cells])
```

### Index above matrix so as so make Group Lasso Matrix Y
```{r eval=FALSE}
ddN_grplasso_Y_vec <- c()

for (DEG in 1:ddN_k) {
  for (cell in 1:ddN_n) {
    ddN_DEG_cell_exp <- ddN_precursor_DEGs_in_precursors_mat[cell, DEG]
    ddN_grplasso_Y_vec <- c(ddN_grplasso_Y_vec, ddN_DEG_cell_exp)
  }
}

# 1665 x 1
ddN_grplasso_Y_mat <- as.matrix(ddN_grplasso_Y_vec, ncol=1, nrow=(ddN_n*ddN_k))
```

### Rename Row names to reflect cell and DEG status
The Row Names follow the same pattern as the Row Names of Group Lasso Matrix X.

```{r eval=FALSE}
ddN_grplasso_Y_DEG_statuses <- rep(c(1:ddN_k), each = ddN_n)
ddN_grplasso_Y_DEGs <- paste0("DEG_", ddN_grplasso_Y_DEG_statuses)
ddN_grplasso_Y_cells <- rep(paste0("_cell_", seq(1, ddN_n)), ddN_k)
ddN_new_grplasso_Y_rownames <- paste0(ddN_grplasso_Y_DEGs, ddN_grplasso_Y_cells)

rownames(ddN_grplasso_Y_mat) <- ddN_new_grplasso_Y_rownames
```

### Final Matrix Y
```{r eval=FALSE}
# 1665 x 1
final_ddN_grplasso_Y_mat <- ddN_grplasso_Y_mat
```

## Run Group Lasso
### Variables: Group Lasso Matrices X and Y and Group IDs
Each TF is a group. Thus, there are p groups each of size k. Given the way Group Lasso Matrix X is organized, the Group IDs follow the pattern of 1, 2, … p; 1, 2, … p; k times. 

```{r eval=FALSE}
grplasso_ddN_groups <- rep(seq(1, ddN_p), ddN_k)

ddN_X <- final_grplasso_X_mat
ddN_Y <- final_ddN_grplasso_Y_mat
ddN_group_id <- grplasso_ddN_groups
```

### Do Cross-Validation
```{r eval=FALSE}
# ~3 mins to Run 
grplasso_ddN_cv_grepreg_fit <- cv.grpreg(ddN_X, ddN_Y, ddN_group_id, penalty="grLasso")
```

### Get Sets of Selected TF Groups at Various Values of Lambda
######## REF SUPP FIGURE PG 31 ##########
```{r eval=FALSE}
# Key: lambda; Values: Set of Selected TF Groups
ddN_group_lassoed_tf_list <- get_group_lassoed_tf_list(grplasso_ddN_cv_grepreg_fit, ddN_candidate_cocktail_tfs)

# Key: lambda; Values: Set of Non-selected TF Groups
ddN_non_group_lassoed_TF_list <- lapply(ddN_group_lassoed_tf_list, get_non_group_lassoed_TFs, ddN_candidate_cocktail_tfs)
```

#### Let us Visualize the Number of Groups Selected at Various Values of Lambda
```{r eval=FALSE}
plot(grplasso_GLRA_cv_grepreg_fit)
```

### Get Sorted Total TF Importance Lists
The Total Importance Score of a Cocktail TF is the sum of the absolute value of its DEG-specific coefficients. It thus gives a quantitative measure of a Cocktail TF's strength over the Gene Regulatory Network (as defined by the Precursor DEGs)
underlying the Precursor's state. 

```{r eval=FALSE}
# (For all lambdas) key = lambda; value is a list with key = TF and value = non0 betas.
ddN_TF_importances_lists <- get_TF_importances_list(grplasso_ddN_cv_grepreg_fit)

# Sum the absolute values of these betas for all TFs for all lambdas.
ddN_sum_total_TF_importance_lists <- ddN_TF_importances_lists %>% map(get_total_summed_TF_importance_list)

# Make sorted dfs given these Total Importance Scores.
ddN_sorted_sum_total_TF_importance_lists <- ddN_sum_total_TF_importance_lists %>% map(make_sorted_TF_importance_df, ddN_candidate_cocktail_tfs)
```

# Final Result: Inferred ddN Cocktail
Upon Inspection of "ddN_group_lassoed_tf_list", we see that the set of 3, the Inferred Cocktail, consists of Dmbx, Lhx1, and HMG1/2.

# Extended Analyses 
>Extended Analyses 1 and 2 compare Cocktail Inference results from another Linear Model, Gaussian Multivariate Multi-Response Linear Regression (GMML), and a Nonlinear Model based on GRNBoost2 respectively to both demonstrate the superiority of the Group Lasso and validate its results. 

> Extended Analysis 3 compares Cocktail Inference results with Terminal ddNs (fully differentiated ddNs) to demonstrate the need to study regulatory changes at the time of specification in order to correctly infer Cocktails.

> Extended Analysis 4 performs subsampling of ddN Precursor data to build confidence for Group Lasso results. 

## (1) Comparison to Gaussian Multivariate Multi-Response Linear Regression (GMML)
DEG expression levels can be modeled as linear combinations of Candidate Cocktail TF expression levels. In our case, we have multiple predictors (TFs) and multiple response variables (expression of DEGs) and both the predictors and response variables are correlated. Thus, an appropriate way to model these interactions is via a gaussian multivariate multi-response linear regression (GMML). 

To infer the Cocktail, we employ 2 strategies:  

Strategy 1: We calculate the average coefficient values for each Candidate
Cocktail TF, get a Total Importance Score (described afterwards), and consider the top 3 based on
the Total Importance Score as the Cocktail.   

Strategy 2: We repeat the same process, but on the **ranks** of the coefficient
values for each Candidate Cocktail TF. 
This is because DEGs with very high expression levels will bias coefficient values
to be higher. We prefer Strategy 2.   

Refer to Methods of paper for a full description of the model.

### Matrices X and Y
Matrix X has expression of Cocktail Candidate TFs in Precursor cells and 
has dimensions n x p. 
Matrix Y has expression of all Precursor DEGs in Precursor cells and
has dimensions n x k.

```{r eval=FALSE}
glm_ddN_X_mat <- as.matrix(ddN_TFs_in_precursor_pseudostate_mat)
glm_ddN_Y_mat <- as.matrix(ddN_precursor_DEGs_in_precursors_mat)
```

### Perform Regression
#### Without Lasso penalty
We will refer to this as the complete GMML model.
```{r eval=FALSE}
# ~Took 5s to run.
ddN_fit <- cv.glmnet(glm_ddN_X_mat, glm_ddN_Y_mat, intercept = FALSE, family = "mgaussian")
ddN_lambda <- ddN_fit$lambda.min
# ddN_coef is a list with keys = khids of Precursor DEGs and values = series of attributes
# one of which (@ x) contains weights for all 14 input TF variables.
ddN_coef_list <- coef(ddN_fit, s = ddN_lambda)
```

#### With Lasso penalty, selecting for 3 TF Variables
Here, we set the dfmax parameter to 2, which applies a Lasso penalty and selects for 3 TF variables.

One can observe the results given different values of this parameter, such as 3 or 4,
which respectively returns 4 or 5 variables, but we study 3 variables in order to 
compare this result to our inferred cocktail, which contains 3 TFs.

```{r eval=FALSE}
# Took ~15s to run.
ddN_fit_3TFs <- cv.glmnet(glm_ddN_X_mat, glm_ddN_Y_mat, intercept = FALSE, family = "mgaussian", dfmax = 2)
ddN_lambda_3TFs <- ddN_fit_3TFs$lambda.min
ddN_coef_3TFs_list <- coef(ddN_fit_3TFs, s = ddN_lambda_3TFs)
```

### Strategy 1
#### Get Regression Coefficients for all Candidate Cocktail TFs
Make a coef_df, which is a data frame with rows = Candidate Cocktail Precursor DEGs, 
columns = Precursor DEGs, and entries as regression coefficients present in coef_list.

```{r eval=FALSE}
ddN_coef_df <- make_coef_df(ddN_coef_list,  ddN_candidate_cocktail_tfs, cleaned_ddN_precursor_pseudostate_DEGs)

ddN_coef_df_3TFs <- make_coef_df(ddN_coef_3TFs_list,  ddN_candidate_cocktail_tfs, cleaned_ddN_precursor_pseudostate_DEGs)
```

### Visualize Distribution of Coefficients and get their Averages
We notice that Dmbx, Lhx1, and HMG1/2 have the highest 1st, 2nd, and 4th average coefficient values in the complete GMML model. 

######## REF SUPP FIGURE PG 32, 33 ##########
```{r eval=FALSE}
ddN_average_coef_df <- get_average_coefficient_df(ddN_coef_df)
ddN_coef_distribution_plot <- get_coefficients_distribution_plot(ddN_coef_df)
ddN_coef_distribution_plot

ddN_average_coef_df_3TFs <- get_average_coefficient_df(ddN_coef_df_3TFs)
```

#### Get Total Importance Scores for Candidate Cocktail TFs
> Sum the absolute values of the regression coefficients

######## REF SUPP FIGURE PG 34 ##########
```{r eval=FALSE}
ddN_total_GMML_importance_df <- get_total_GMML_importance_df(ddN_coef_df)
ddN_total_GMML_importance_df_3TFs <- get_total_GMML_importance_df(ddN_coef_df_3TFs)
```

### Strategy 2
#### First Get the ranks of the absolute values of the regression coefficients

```{r eval=FALSE}
ranked_ddN_coef_df <- ddN_coef_df %>% mutate(across(everything(), abs), across(everything(), min_rank))
ranked_ddN_coef_df_3TFs <- ddN_coef_df_3TFs %>% mutate(across(everything(), abs), across(everything(), min_rank))
```

#### Visualize Distribution of Coefficients and get their Averages
We notice that Dmbx, Lhx1, and HMG1/2 have the highest 1st, 2nd, and 4th average ranked coefficient values in the complete GMML model. 

######## REF SUPP FIGURE PG 32, 34 ##########
```{r eval=FALSE}
ranked_ddN_average_coef_df <- get_average_coefficient_df(ranked_ddN_coef_df)
ranked_ddN_coef_distribution_plot <- get_coefficients_distribution_plot(ranked_ddN_coef_df)
ranked_ddN_coef_distribution_plot

ranked_ddN_average_coef_df_3TFs <- get_average_coefficient_df(ranked_ddN_coef_df_3TFs)
```
#### Get Total Importance Scores for Candidate Cocktail TFs

######## REF SUPP FIGURE PG 35 ##########
```{r eval=FALSE}
ranked_ddN_total_GMML_importance_df <- get_total_GMML_importance_df(ranked_ddN_coef_df)

ranked_ddN_total_GMML_importance_df_3TFs <- get_total_GMML_importance_df(ranked_ddN_coef_df_3TFs)
```

### Notes on Final Result
> In this case, given the above results, the ddN Cocktail is likely inferred from the GMML model.

## (2) Comparison with GRNBoost2 Inferences
The purpose of this part of the analysis is to compare Cocktail predictions derived
from nonlinear models (GRNBoost2) with those from the linear Group Lasso model, 
as TF-gene relationships may be better modeled as nonlinear ones.

This Pipeline returns 2 lists:
sorted_total_importance_tfs and sorted_percent_top_list, which are used in tandem
to infer the Cocktail. Please refer to documentation in "GRNBoost2_pipeline.py"
to see the methods behind deriving the lists.

```{r eval=FALSE}
ddN_GRN_dir <- "GRN/ddN/"
```

#### ddN Precursor DEG Counts Matrix
```{r eval=FALSE}
# Matrix must have counts from both DEGs and Candidate Cocktail TFs.
ddN_precursor_pseudostate_DEGs_plus <- unique(c(ddN_precursor_pseudostate_DEGs, ddN_candidate_cocktail_khids))
ddN_precursor_pseudostate_GRN_mat <- URD_raw_data[ddN_precursor_pseudostate_DEGs_plus, ddN_precursor_pseudostate_cells]
# 9 x 210
ddN_precursor_pseudostate_GRN_df <- as.data.frame(t(ddN_precursor_pseudostate_GRN_mat))
```

#### Inputs to GRNBoost2 Pipeline
```{r eval=FALSE}
write_tsv(ddN_precursor_pseudostate_GRN_df, paste0("GRN/", "ddN_precursor_pseudostate_GRN_df.tsv"))
```

```{r eval=FALSE}
write_csv_helper(ddN_candidate_cocktail_khids, paste0("GRN/", "ddN_candidate_cocktail_khids.csv"))
```

#### Do GRNBoost2 50 times.
```{r eval=FALSE}
source_python("GRNBoost2_pipeline.py")
```

```{r eval=FALSE}
set.seed(16)

expression_count_matrix_tsv <- paste0("GRN/", "ddN_precursor_pseudostate_GRN_df.tsv")
tf_khid_csv <- paste0("GRN/", "ddN_candidate_cocktail_khids.csv")
outname <- "GRN/ddN/ddN"
min_i <- 0L
max_i <- 50L

# ~Took 15 mins to run.
do_GRNBoost2(expression_count_matrix_tsv, tf_khid_csv, outname, min_i, max_i)
```

##### sorted_total_importance_tfs
Each element of the list is a tuple of TF and list with Mean and Standard Deviation
of Total Importance Scores. The TFs are sorted in nonincreasing order based on the mean.
           
##### sorted_percent_top_list
Each element of the list is a TF and its Percentage Covered as a Top 3 Regulator for all DEGs.

```{r eval=FALSE}
khid_tf_tsv_file <- "/Users/prakritipaul/Git/Cocktails/useful_lists/kai_tf.tsv"
ddN_GRN_path <- "/Users/prakritipaul/Git/Cocktails/FINAL_code_and_results/GRN/ddN/*.tsv"
n <- 3L

# Returns sorted_total_importance_tfs and sorted_percent_top_list
ddN_GRNBoost2_outputs <- GRNBoost2_pipeline(ddN_GRN_path, khid_tf_tsv_file, n)
```

##### Inspect Output contents.
```{r eval=FALSE}
ddN_sorted_total_importance_tfs <- ddN_GRNBoost2_outputs[1]
ddN_sorted_percent_top_list <- ddN_GRNBoost2_outputs[2]
```

### Notes on Final Result
> Upon inspection of the sorted_total_importance_tfs and sorted_percent_top-list, 
only Dmbx and Lhx1 appear in the Top 3. In this case, this method is not sufficient. However, this result does further support the results obtained from the Group Lasso Model. 

## (3) Comparison with Method Applied on Terminal ddNs
A necessary comparison is with Terminal ddNs, or fully differentiated ddNs, as 
a key claim in this work is that the Cocktail cannot be inferred without 
studying cell states at the time of specification. 

Potentially, we can find the Cocktail either in the top 50% mostly highly expressed DEG TFs in the Terminal DEGs (these are Terminal Candidate Cocktail TFs) or in the output of running the Group Lasso Model using Terminal DEGs and Terminal Candidate Cocktail Factors.
We will also use the GMML.

From the following analyses, we demonstrate that the Cocktail cannot be inferred from either way and build evidence for the power of our computational method and the strength of the assumptions we made.

### Identify Terminal ddNs and their Siblings
Siblings were identified from observing the Virtual Lineage Tree.

```{r eval=FALSE}
# 64
terminal_GLRAs <- neural_CT_list$`GLRA1.2.3+ motor ganglion`
# 18
terminal_ddNs <- get_dmbx_expressing_cells(terminal_GLRAs, pseudotimes)
# 117
terminal_ddN_siblings <- c(neural_CT_list$`KCNB1+ motor ganglion`, neural_CT_list$`AMD+ motor ganglion` )
```

### Raw Mats
```{r eval=FALSE}
terminal_ddN_raw_mat <- URD_raw_data[, terminal_ddNs]
terminal_ddN_TF_raw_mat <- URD_raw_data[tfs_in_URD, terminal_ddNs]

terminal_ddN_sibling_raw_mat <- URD_raw_data[, terminal_ddN_siblings]
terminal_ddN_sibling_TF_raw_mat <- URD_raw_data[tfs_in_URD, terminal_ddN_siblings]
```

### Calculate Terminal DEGs with respect to Siblings
Terminal DEGs are analogous to Precursor DEGs. 

We calculate Terminal DEG TFs with respect to Siblings because comparison to all other neural
cell types did not render meaningful TFs. Instead, TFs known to play roles in specification appeared as differentially expressed when Terminal ddNs were compared to its siblings. This made these results more competitive with the results obtained from applying the computational method to Precursors at the time of specification.

### Get Terminal DEGs
```{r eval=FALSE}
terminal_ddN_v_siblings_mat <- as.matrix(cbind(terminal_ddN_raw_mat, terminal_ddN_sibling_raw_mat))

terminal_ddN_v_siblings_edgeR_outputs <- complete_edgeR_pipeline(terminal_ddN_v_siblings_mat, ncol(terminal_ddN_raw_mat), ncol(terminal_ddN_sibling_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 297
terminal_ddN_v_siblings_DEGs_df <- terminal_ddN_v_siblings_edgeR_outputs[[1]]
terminal_ddN_v_siblings_DEGs <- rownames(terminal_ddN_v_siblings_DEGs_df)
```

#### Write them out.
```{r eval=FALSE}
write_csv_helper(terminal_ddN_v_siblings_DEGs, paste0(DEG_lists_out_dir, "terminal_ddN_v_siblings_DEGs.csv"))
```

### Get Terminal DEG TFs
We will filter these downstream to get Terminal Candidate Cocktail TFs.
```{r eval=FALSE}
terminal_ddN_v_siblings_TF_mat <- as.matrix(cbind(terminal_ddN_TF_raw_mat, terminal_ddN_sibling_TF_raw_mat))

terminal_ddN_v_siblings_TF_edgeR_outputs <- complete_edgeR_pipeline(terminal_ddN_v_siblings_TF_mat, ncol(terminal_ddN_TF_raw_mat), ncol(terminal_ddN_sibling_TF_raw_mat), 0.05, -1, 50)
```

```{r eval=FALSE}
# 14
terminal_ddN_v_siblings_TF_DEGs_df <- terminal_ddN_v_siblings_TF_edgeR_outputs[[1]]
# Get TF orthologs.
named_terminal_ddN_v_siblings_TF_DEGs_df <- convert_khids_to_ortholog(terminal_ddN_v_siblings_TF_DEGs_df)
terminal_ddN_v_sibling_TF_DEGs <- named_terminal_ddN_v_siblings_TF_DEGs_df$Ghost.Name
```

#### Write them out.
```{r eval=FALSE}
write_csv_helper(terminal_ddN_v_sibling_TF_DEGs, paste0(DEG_lists_out_dir, "terminal_ddN_v_sibling_TF_DEGs.csv"))
```

### Get Terminal Candidate Cocktail TFs
As before, we will take the top 50% most highly expressed of the Terminal DEG TFs and consider them as the Terminal Candidate Cocktail TFs.

######## REF SUPP FIGURE PG 35 ##########
```{r eval=FALSE}
# Get mean Expression of all DEG TFs
terminal_ddN_mean_TF_exp_df <- make_named_mean_TF_exp_df(terminal_ddNs)

# Get khids of Terminal DEG TFs
terminal_ddN_v_siblings_TF_DEG_khids <- rownames(terminal_ddN_v_siblings_TF_DEGs_df)

# Get Mean expression of Terminal DEG TFs in terminal cells.
terminal_ddN_cocktail_TF_mean_exp_df <- terminal_ddN_mean_TF_exp_df %>% filter(KH.gene.model %in% terminal_ddN_v_siblings_TF_DEG_khids)

# Get top 50% most highly expressed of above DEG TFS
terminal_cutoff <- 0.5

terminal_ddN_cutoff_df <- terminal_ddN_cocktail_TF_mean_exp_df[1:(terminal_cutoff*(length(terminal_ddN_v_siblings_TF_DEG_khids))), ]

# Get the TFs and khids
# 7
terminal_ddN_candidate_cocktail_khids <- terminal_ddN_cutoff_df$KH.gene.model
terminal_ddN_candidate_cocktail_tfs <- terminal_ddN_cutoff_df$Ghost.Name
```

#### Write out the Candidate Cocktail TFs
```{r eval=FALSE}
write_csv_helper(terminal_ddN_candidate_cocktail_tfs, paste0(DEG_lists_out_dir, "terminal_ddN_candidate_cocktail_tfs.csv"))
```

#### Brief Aside
An interesting observation is that the common Candidate Cocktail TFs between the Precursor and Terminal ddN are the Cocktail itself except for HMG1/2 (Pioneer factor).
```{r eval=FALSE}
common_ddN_cocktail_TFs <- intersect(terminal_ddN_candidate_cocktail_tfs, ddN_candidate_cocktail_tfs)
common_ddN_cocktail_TFs
```

### (a) Do Cocktail TF Inference using Group Lasso Model
#### Variables
```{r eval=FALSE}
# Get rid of TFs in DEGs.
# 7
common_terminal_ddN_DEG_TFs <- intersect(terminal_ddN_candidate_cocktail_khids, terminal_ddN_v_siblings_DEGs)
# 290
cleaned_terminal_ddN_DEGs <- setdiff(terminal_ddN_v_siblings_DEGs, common_terminal_ddN_DEG_TFs)
```

```{r eval=FALSE}
# 18, 7, 290
terminal_ddN_n <- length(terminal_ddNs)
terminal_ddN_p <- length(terminal_ddN_candidate_cocktail_tfs)
terminal_ddN_k <- length(cleaned_terminal_ddN_DEGs)
```

#### Make Group Lasso X Matrix
```{r eval=FALSE}
# 18 x 7
terminal_ddN_TFs_in_terminal_mat <- t(URD_data[terminal_ddN_candidate_cocktail_khids, terminal_ddNs])
```

```{r eval=FALSE}
# 5220 x 2030
terminal_ddN_grplasso_X_mat <- make_grplasso_X_mat(terminal_ddN_TFs_in_terminal_mat, terminal_ddN_n, terminal_ddN_p, terminal_ddN_k)
```

#### Make Group Lasso Y Matrix
```{r eval=FALSE}
# 18 x 290
terminal_ddN_DEGs_in_terminal_mat <- t(URD_data[cleaned_terminal_ddN_DEGs, terminal_ddNs])
```

```{r eval=FALSE}
# 5220    1
terminal_ddN_grplasso_Y_mat <- make_grplasso_Y_mat(terminal_ddN_DEGs_in_terminal_mat, terminal_ddNs, terminal_ddN_p, terminal_ddN_k)
```

#### Variables: Group Lasso Matrices X and Y and Group IDs
```{r eval=FALSE}
terminal_grplasso_ddN_groups <- rep(seq(1, terminal_ddN_p), terminal_ddN_k)

terminal_ddN_X <- terminal_ddN_grplasso_X_mat
terminal_ddN_Y <- terminal_ddN_grplasso_Y_mat
terminal_ddN_group_id <- terminal_grplasso_ddN_groups
```

#### Error: Group Lasso does not solve. 
```{r eval=FALSE}
# 
terminal_grplasso_ddN_cv_grepreg_fit <- cv.grpreg(terminal_ddN_X, terminal_ddN_Y, terminal_ddN_group_id, penalty="grLasso")
```

### Get Sets of Selected TF Groups at Various Values of Lambda
```{r eval=FALSE}
# Key: lambda; Values: Set of Selected TF Groups
terminal_ddN_group_lassoed_tf_list <- get_group_lassoed_tf_list(terminal_grplasso_ddN_cv_grepreg_fit, terminal_ddN_candidate_cocktail_tfs)

# Key: lambda; Values: Set of Non-selected TF Groups
terminal_ddN_non_group_lassoed_TF_list <- lapply(terminal_ddN_group_lassoed_tf_list, get_non_group_lassoed_TFs, terminal_ddN_candidate_cocktail_tfs)
```

#### Let us Visualize the Number of Groups Selected at Various Values of Lambda
```{r eval=FALSE}
plot(terminal_grplasso_ddN_cv_grepreg_fit)
```

### Get Sorted Total TF Importance Lists
The Total Importance Scores within these lists give us a quantitative measure of a 
Cocktail TF's strength over the Gene Regulatory Network (as defined by the Precursor DEGs)
underlying the Precursor's state. 

```{r eval=FALSE}
# (For all lambdas) key = lambda; value is a list with key = TF_num and value = betas.
terminal_ddN_TF_importances_lists <- get_TF_importances_list(terminal_grplasso_ddN_cv_grepreg_fit)

# Sum the absolute values of these betas for all TFs for all lambdas.
terminal_ddN_sum_total_TF_importance_lists <- terminal_ddN_TF_importances_lists %>% map(get_total_summed_TF_importance_list)

# Make sorted dfs given these Total Importance Scores.
terminal_ddN_sorted_sum_total_TF_importance_lists <- terminal_ddN_sum_total_TF_importance_lists %>% map(make_sorted_TF_importance_df, terminal_ddN_candidate_cocktail_tfs)
```

#### Result
The model does not converge. We cannot know if the Group Lasso was successful.

### (b) Do GMML 
```{r eval=FALSE}
glm_terminal_ddN_X_mat <- as.matrix(terminal_ddN_TFs_in_terminal_mat)
glm_terminal_ddN_Y_mat <- as.matrix(terminal_ddN_DEGs_in_terminal_mat)
```

### Perform Regression
#### Without Lasso penalty
We will refer to this as the complete GMML model.
```{r eval=FALSE}
# ~Took 40s to run.
terminal_ddN_fit <- cv.glmnet(glm_terminal_ddN_X_mat, glm_terminal_ddN_Y_mat, intercept = FALSE, family = "mgaussian")
terminal_ddN_lambda <- terminal_ddN_fit$lambda.min
# terminal_ddN_coef is a list with keys = khids of Precursor DEGs and values = series of attributes
# one of which (@ x) contains weights for all 14 input TF variables.
terminal_ddN_coef_list <- coef(terminal_ddN_fit, s = terminal_ddN_lambda)
```

#### With Lasso penalty, selecting for 3 TF Variables
Here, we set the dfmax parameter to 2, which applies a Lasso penalty and selects for 3 TF variables.

One can observe the results given different values of this parameter, such as 3 or 4,
which respectively returns 4 or 5 variables, but we study 3 variables in order to 
compare this result to our inferred cocktail, which contains 3 TFs.

```{r eval=FALSE}
# Took ~30s to run.
terminal_ddN_fit_3TFs <- cv.glmnet(glm_terminal_ddN_X_mat, glm_terminal_ddN_Y_mat, intercept = FALSE, family = "mgaussian", dfmax = 2)
terminal_ddN_lambda_3TFs <- terminal_ddN_fit_3TFs$lambda.min
terminal_ddN_coef_3TFs_list <- coef(terminal_ddN_fit_3TFs, s = terminal_ddN_lambda_3TFs)
```

### Strategy 1
#### Get Regression Coefficients for all Candidate Cocktail TFs
Make a coef_df, which is a data frame with rows = Candidate Cocktail Precursor DEGs, 
columns = Precursor DEGs, and entries as regression coefficients present in coef_list.

```{r eval=FALSE}
terminal_ddN_coef_df <- make_coef_df(terminal_ddN_coef_list,  terminal_ddN_candidate_cocktail_tfs, cleaned_terminal_ddN_DEGs)

terminal_ddN_coef_df_3TFs <- make_coef_df(terminal_ddN_coef_3TFs_list,  terminal_ddN_candidate_cocktail_tfs, cleaned_terminal_ddN_DEGs)
```

#### Visualize Distribution of Coefficients and get their Averages
Only Dmbx has a solved coefficient.

######## REF SUPP FIGURE PG 36, 37 ##########
```{r eval=FALSE}
terminal_ddN_average_coef_df <- get_average_coefficient_df(terminal_ddN_coef_df)
terminal_ddN_coef_distribution_plot <- get_coefficients_distribution_plot(terminal_ddN_coef_df)
terminal_ddN_coef_distribution_plot

terminal_ddN_average_coef_df_3TFs <- get_average_coefficient_df(terminal_ddN_coef_df_3TFs)
```

#### Get Total Importance Scores for Candidate Cocktail TFs
######## REF SUPP FIGURE PG 38 ##########
```{r eval=FALSE}
terminal_terminal_ddN_total_GMML_importance_df <- get_total_GMML_importance_df(terminal_ddN_coef_df)

terminal_ddN_total_GMML_importance_df_3TFs <- get_total_GMML_importance_df(terminal_ddN_coef_df_3TFs)
```

### Strategy 2
#### First Get the ranks of the absolute values of the regression coefficients
```{r eval=FALSE}
ranked_terminal_ddN_coef_df <- terminal_ddN_coef_df %>% mutate(across(everything(), abs), across(everything(), min_rank))

ranked_terminal_ddN_coef_df_3TFs <- terminal_ddN_coef_df_3TFs %>% mutate(across(everything(), abs), across(everything(), min_rank))
```

#### Visualize Distribution of Coefficients and get their Averages
Only Dmbx has a solved coefficient.

######## REF SUPP FIGURE PG 37 ##########
```{r eval=FALSE}
ranked_terminal_ddN_average_coef_df <- get_average_coefficient_df(ranked_terminal_ddN_coef_df)

ranked_terminal_ddN_coef_distribution_plot <- get_ddN_coefficients_distribution_plot(ranked_terminal_ddN_coef_df)
ranked_terminal_ddN_coef_distribution_plot

ranked_terminal_ddN_average_coef_df_3TFs <- get_average_coefficient_df(ranked_terminal_ddN_coef_df_3TFs)
```

#### Get Total Importance Scores for Candidate Cocktail TFs
######## REF SUPP FIGURE PG 38 ##########
```{r eval=FALSE}
ranked_terminal_ddN_total_GMML_importance_df <- get_total_GMML_importance_df(ranked_terminal_ddN_coef_df)

ranked_terminal_ddN_total_GMML_importance_df_3TFs <- get_total_GMML_importance_df(ranked_terminal_ddN_coef_df_3TFs)
```


### (c) Comparison with GRNBoost2 Inferences
The purpose of this part of the analysis is to compare Cocktail predictions derived
from nonlinear models (GRNBoost2) with those from the linear Group Lasso model, 
as TF-gene relationships may be better modeled as nonlinear ones.

This Pipeline returns 2 lists:
sorted_total_importance_tfs and sorted_percent_top_list, which are used in tandem
to infer the Cocktail. Please refer to documentation in "GRNBoost2_pipeline.py"
to see the methods behind deriving the lists.

```{r eval=FALSE}
terminal_ddN_GRN_dir <- "./GRN/terminal_ddN/"
```

##### Terminal ddN DEG Counts Matrix
```{r eval=FALSE}
# Matrix must have counts from both DEGs and Candidate Cocktail TFs.
terminal_ddN_DEGs_plus <- unique(c(terminal_ddN_v_siblings_DEGs, terminal_ddN_candidate_cocktail_khids))
terminal_ddN_GRN_mat <- URD_raw_data[terminal_ddN_DEGs_plus, terminal_ddNs]

# 18 x 297
terminal_ddN_GRN_df <- as.data.frame(t(terminal_ddN_GRN_mat))
```

#### Inputs to GRNBoost2 Pipeline
```{r eval=FALSE}
write_tsv(terminal_ddN_GRN_df, paste0("GRN/", "terminal_ddN_GRN_df.tsv"))
```

```{r eval=FALSE}
write_csv_helper(terminal_ddN_candidate_cocktail_khids, paste0("GRN/", "terminal_ddN_candidate_cocktail_khids.csv"))
```

#### Do GRNBoost2 50 times.
```{r eval=FALSE}
set.seed(16)

expression_count_matrix_tsv <- paste0("GRN/", "terminal_ddN_GRN_df.tsv")
tf_khid_csv <- paste0("GRN/", "terminal_ddN_candidate_cocktail_khids.csv")
outname <- "GRN/terminal_ddN/terminal_ddN"
min_i <- 0L
max_i <- 50L

# Took ~9 mins to run.
do_GRNBoost2(expression_count_matrix_tsv, tf_khid_csv, outname, min_i, max_i)
```

##### sorted_total_importance_tfs
Each element of the list is a tuple of TF and list with Mean and Standard Deviation
of Total Importance Scores. The TFs are sorted in nonincreasing order based on the mean.
           
##### sorted_percent_top_list
Each element of the list is a TF and its Percentage Covered as a Top 3 Regulator for all DEGs.

```{r eval=FALSE}
khid_tf_tsv_file <- "./helper_files/Ciona_khid_TF.tsv"
terminal_ddN_GRN_path <- "/Users/prakritipaul/Git/Cocktails/FINAL_code_and_results/GRN/terminal_ddN/*.tsv"
n <- 3L

# Returns sorted_total_importance_tfs and sorted_percent_top_list
terminal_ddN_GRNBoost2_outputs <- GRNBoost2_pipeline(terminal_ddN_GRN_path, khid_tf_tsv_file, n)
```

##### Inspect Output contents.
```{r eval=FALSE}
terminal_ddN_sorted_total_importance_tfs <- terminal_ddN_GRNBoost2_outputs[1]
terminal_ddN_sorted_percent_top_list <- terminal_ddN_GRNBoost2_outputs[2]
```

#### Note on Final Result
##### We are once again able to uncover the Cocktail from studying the terminal state. This further supports the need to study gene regulatory changes occurring at the time of specification.

## (4) Subsampling
> Finally, in order to build evidence for our predictions, we sample 8 cells 9 times, 
run the Group Lasso Model, and keep track of its inferred Cocktail. 
We count the number of times that a TF in our predicted Cocktail appears in a subsample's inferred Cocktail and thus get a percentage/confidence for the overall prediction. 

>In our analyses, we obtain a 100% confidence for each of the TFs Dmbx, Lhx1, and HMG1/2 in the ddN Cocktail.

### Subsample 8 cells 9 times. 
```{r eval=FALSE}
subsampling_out_dir <- "/Users/prakritipaul/Git/Cocktails/FINAL_code_and_results/subsampling/"
```

# Do subsampling by leaving one cell out.
```{r eval=FALSE}
ddN_seed <- 8
# This list contains output of grepreg_routine for all 9 rounds of subsampling.
ddN_subsampling_output_list <- list()

i <- 1
for (ddN_cell in ddN_precursor_pseudostate_cells) {
  cat("Doing for i = ", i, "cell = ", ddN_cell, "\n")
  ddN_cells <- ddN_precursor_pseudostate_cells[!ddN_precursor_pseudostate_cells %in% ddN_cell]
  
  # Sometimes the model doesn't converge.
  tryCatch({ddN_subsampling_output <- grepreg_routine(ddN_cells,
                                  ddN_candidate_cocktail_tfs,
                                  cleaned_ddN_precursor_pseudostate_DEGs,
                                  ddN_TFs_in_precursor_pseudostate_mat,
                                  ddN_precursor_DEGs_in_precursors_mat,
                                  ddN_seed)
  ddN_subsampling_output_list[[i]] <- ddN_subsampling_output
  
  }, error = function(error) {
    cat(paste("An error occurred for ddN cell", ddN_cell, "and i =", i, "\n"))
    ddN_subsampling_output_list[[i]] <- ddN_cell
  })
  i <- i+1
}
```

# Let's make the subsampled group tf df.
```{r eval=FALSE}
ddN_combined_subsampled_grouped_tf_df <- make_lambda_to_grouped_tf_df(ddN_subsampling_output_list)
```

# Write it out.
```{r eval=FALSE}
write_out_group_lasso_tf_df(ddN_combined_subsampled_grouped_tf_df, subsampling_out_dir, "ddN_combined_subsampled_tf_df.csv")
```

# Also write out CT_group_lassoed_tf_list.
```{r eval=FALSE}
write_out_CT_group_lassoed_tf_list(ddN_group_lassoed_tf_list, subsampling_out_dir, "ddN_group_lassoed_tf_df.csv")
```

### Run Subsampling Prediction Confidence Pipeline
```{r eval=FALSE}
source_python("subsampling_prediction_confidence.py")
```

```{r eval=FALSE}
ddN_group_lassoed_tf_df_csv <- paste0(subsampling_out_dir, "ddN_group_lassoed_tf_df.csv")
ddN_subsampled_tf_df_csv <- paste0(subsampling_out_dir, "ddN_combined_subsampled_tf_df.csv")
ddN_tf_set_sizes <- list(3, 4, 5, 6)
ddN_ideal_set_size <- 3

ddN_cocktail_subsampling_confidence <- subsampling_prediction_confidence(ddN_group_lassoed_tf_df_csv, ddN_subsampled_tf_df_csv, ddN_tf_set_sizes, ddN_ideal_set_size)

ddN_cocktail_subsampling_confidence
```
# RSession Info
> Refer to Supplement for complete information.

```{r include=TRUE}
sessionInfo()
```

